Cyclic muscle contractions reinforce the acto-myosin motors and mediate the full elongation of C. elegans embryo

Version 1

Abstract

Abstract

The paramount importance of mechanical forces in morphogenesis and embryogenesis is widely recognized, but understanding the mechanism at the cellular and molecular level is still challenging. Due to its simple internal organization, Caenorhabditis elegans is a worthwhile study system. As demonstrated experimentally, after a first period of steady elongation due to the acto-myosin network, muscle contractions operate a quasi-periodic sequence of bending, rotation and torsion, which leads to the final 4-fold size of the embryo before hatching. How acto-myosin and muscles contribute to embryonic elongation is herein theoretically studied. A filamentary elastic model that converts stimuli generated by biochemical signals in the tissue into driving forces, explains embryonic deformation under actin bundles and muscle activities, and dictates mechanisms of late elongation based on the effects of energy conversion and dissipation. We quantify this dynamic transformation by stretches applied to a cylindrical structure mimicking the body shape in finite elasticity, obtaining good agreement and understanding for both wild-type and mutant embryos at all stages.

Introduction

Mechanical stresses play a critical role during embryogenesis, affecting complex biological tissues such as the skin, the brain, and the interior of organs. There are many studies in the last decades, for example, the pattern of the intestine (Coulombre and Coulombre (1958); Hannezo et al. (2011); Li et al. (2011); Ben Amar and Jia (2013); Shyer et al. (2013)), the brain cortex (Toro and Burnod (2005); Goriely et al. (2015); Ben Amar and Bordner (2017); Tallinen et al. (2016)), and the cir-cumvolutions of the fingerprints (Kücken and Newell (2005); Ciarletta and Ben Amar (2012)) that have been interpreted as the result of compressive stresses generated by growth occurring a few months post fertilization for humans. At a smaller scale, the influence of mechanical stresses superposed to cellular processes (e.g., division, migration) and tissue organization becomes more complex to be identified and quantified. For example, independently of growth, stresses may be generated by different molecular motors, among which myosin II, linked to actin filaments, is the most prevalent during epithelial cell morphogenesis (Vicente-Manzanares et al. (2009)) and cell motility (Cowan and Hyman (2007); Olson and Sahai (2009); Palumbo et al. (2022)). The spatial distribution and dynamics of myosin II greatly influence the morphogenetic process (Levayer and Lecuit (2012); Lv et al. (2022)), as demonstrated for Drosophila (Bertet et al. (2004); Blankenship et al. (2006); Saxena et al. (2014); Shindo and Wallingford (2014)) and also C. elegans embryo (Priess and Hirsh (1986); Gally et al. (2009); Ben Amar et al. (2018)).

The embryonic elongation of C. elegans represents an attractive model of matter reorganization without a mass increase before hatching. It happens after the ventral enclosure and takes around 240 minutes to convert the lima-bean-shaped embryo into a final elongated worm shape: the embryo elongates along the anterior/posterior axis by four times (McKeown et al. (1998)). The short lifetime of the egg before hatching and its transparency makes this system an ideal system to study the forces that exist in the cortical epithelium or its vicinity. However, unlike the embryonic development of Drosophila and Zebrafish, there is neither cell migration, cell division nor a no-table change in embryonic volume (Sulston et al. (1983); Priess and Hirsh (1986)), only a noticeable epidermal elongation drives the whole morphogenetic process of C. elegans in the period post en-closure. There are two experimentally identified driving forces for the elongation: the actomyosin contractility in seam cells of the epidermis, which seems to last during the whole elongation process, and the muscle activity beneath the epidermis that starts after the 1.7 ∼ 1.8-fold stage. The transition is well defined since the muscle participation makes the embryo rather motile impeding any physical experiments such as laser ablation, which could be conducted and achieved in the first period (Vuong-Brender et al. (2017a)). Consequently, the elongation process of C. elegans could be divided into two stages: the early elongation and the late elongation, depending on the muscle activation, and Fig.1(A) shows the whole elongation process (Vuong-Brender et al. (2017a)). Previously, the role of the actomyosin network in the seam cells during the early elongation of C. elegans was investigated (Ben Amar et al. (2018)). Based on the geometry of a hollow cylinder composed of four parts (seam and dorso-ventral cells), a model involving the pre-stress responsible for the enclosure, the active compressive ortho-radial stress joined to the passive stress quantitatively predicts the elongation, but only up to ∼ 70% of the initial length.

(A) Overview of C. elegans embryonic elongation. Three epidermal cell types are found around the circumference: dorsal, ventral and lateral. (B) Schemes showing a C. elegans cross-section of the embryo. Small yellow arrows in the left image indicate the contraction force that occurred in the seam cell. Four muscle bands underneath the epidermis and actin bundles surround the outer epidermis.

For the late elongation phase, the acto-myosin network and muscles drive together the full elongation during this phase (Vuong-Brender et al. (2017a)). Muscles play an important role since mutants with muscle defects are unable to complete the elongation process, even though the actomyosin network operates normally (Lardennois et al. (2019)). Fig.1(B) shows the schematic image of C. elegans body (Zhang et al. (2011)) with four rows of muscles, two of which are underneath the dorsal epidermis and the other two are under the ventral epidermis. As observed in vivo, C. elegans exhibits systematic rotations accompanying each contraction (Yang (2017); Yang Xinyi et al. (2023)), and deformations such as bending and twisting. Eventually, the C. elegans embryo will complete an elongation from 1.8-fold to 4-fold. We can imagine that once the muscle is activated on one side, it can only contract, and then the contraction forces will be transmitted to the epidermis on this side. So one can wonder how an elongation can occur since striated muscles can only perform cycles of contraction and relaxation, their action will tend to reduce the length of the embryo. Therefore, it is necessary to understand how the embryo elongates during each contraction and how the muscle contractions couple to the acto-myosin activity. This work aims to answer this paradox within the framework of finite elasticity without invoking cell plasticity and stochasticity which cannot be considered driving forces. In addition, several important issues at this stage remain unsettled. First of all, we may also observe a torsion of the embryo but if the muscle activity suggests a bending, it cannot explain alone a substantial torsion. Secondly, a small deviation of the muscle axis (Moerman and Williams (2006)) is responsible for a series of rotations, how to relate these rotations to the muscle activation (Yang (2017)). Since any measurement on a motile embryo at this scale is difficult, it is meaningful to explore the mechanism of late elongation theoretically. Furthermore, muscle contraction is crucial in both biological development and activities and has been studied extensively by researchers (Tan and De Vita (2015); De Vita et al. (2017)), but how it works at small scales remains a challenge.

Using a finite elasticity model and assuming the embryonic body shape is cylindrical, we can evaluate the geometric bending deformation and the energy released during each muscular contraction on one side since after each contraction, the muscles relax and then muscles on the opposite side undergo a new contraction. This cyclic process leads to a tiny elongation of the cylinder along its symmetry axis at each contraction. Each of them is correlated with the rotational movements of the embryo (Yang (2017)). By repeating these pairs of contractions more than two hundred times, a cumulative extension is achieved, but it must be reduced by friction mechanisms, also evaluated by the model. Furthermore, the mechanical model explains the existence of a torque operating at the position of the head or tail by the coupling of muscle contraction with the orthoradial acto-myosin forces. Finally, the small deviation between the muscles and the central axis experimentally detected (Moerman and Williams (2006)) induces cyclic rotations and possibly torsion leading to fluid viscous flow inside the egg. Quantifying all these processes allows evaluating the physical quantities of the embryo such as the shear modulus of each component, the osmolarity of the interstitial fluid and the active forces exerted by the acto-myosin network and the muscles which are sparsely known in the embryonic stages.

Our mechanical model accounts for the dynamical deformations induced by the internal stimuli of a layered soft cylinder, enabling us to make accurate quantitative predictions about the active networks of the embryo. Furthermore, our model takes into consideration the dissipation that occurs in the late period just before the egg hatch. Our results are consistent with observations of acto-myosin and muscle activity (Vuong-Brender et al. (2017a); Ben Amar et al. (2018); Lardennois et al. (2019)). The architecture of our paper is illustrated in Fig.2.

Architecture of the program. The program reflects the framework of research. On the one hand, the proposed model explains the early and late elongation of the C. elegans, on the other hand, the early myosin activity is estimated, the deformations (bending, twisting, rotation) happening in the late period is recovered, and the estimation of energy dissipated during muscle activity is achieved.

Results

It is widely recognized that C. elegans is a well-established model organism in the field of developmental biology. However, it is less widely known that its internal striated muscles share similarities with skeletal muscles found in vertebrates in terms of both function and structure (Lesanpezeshki et al. (2021)). Being contractile, the role of the four axial muscles (as shown in Fig.1) in the final shaping of the embryo is nearly contra-intuitive. In this section, we aim to highlight the action of these muscles when coupled to the acto-myosin network, using a purely mechanical approach. To achieve a quantitative understanding, our method requires a detailed description of the deformation geometry of the embryo, with the body shape represented by a fully heterogeneous cylinder.

Geometric and strain deformations of the embryo

The early elongation of C. elegans embryo was previously studied, which is characterized by an inner cylinder surrounded by epithelial cells located in the cortical position (Ben Amar et al. (2018)). The cortex is composed of three distinct cells - the seam, dorsal, and ventral cells - which exhibit unique cytoskeleton organization and actin network configurations. Among these, only the seam cells possess active myosin motors that function in the ortho-radial direction, allowing for the contraction of the circumference and thereby triggering the early elongation process. In this study, we simplify the cylindrical geometry of the body and treat the epidermis as a unified whole with effective activity localized along the circumference, supported by four muscles distributed beneath the epidermis. Muscles play no role in the early elongation stage so were not considered previously (Vuong-Brender et al. (2017a); Ben Amar et al. (2018)). After this period of approximately a 1.8-fold increase in length, muscles parallel to the main axis and actin bundles organized along the circumference will collaborate to facilitate further elongation. Table. 1 provides the size parameters of C. elegans which will be introduced hereafter.

We focus on the overall deformation of a full cylinder or a thin rod, having a length L greater than the radius R and a central vertical axis along the Z direction. Contrary to previous works (Ciarletta et al. (2009); Ben Amar et al. (2018)), here we decide to simplify the geometrical aspect because of the mechanical complexity. The biological activity induced by the acto-myosin network and muscles is represented by active strains, and the global shape is the result of the coupling between elastic and active strains, modulated by dissipation. Active strains are generated through non-mechanical processes (e.g., biochemical processes): myosin motors and striated muscles receive their energy from ATP hydrolysis, which is converted to mechanical contractions of fibers. Due to the significant deformation observed, the central line is distorted and becomes a curve in three-dimensional space, represented by a vector r (Z), as depicted in Fig.3(A). Along this curve, perpendicular planar sections of the embryo can be defined, and the deformation in each section can be quantified since the circular geometry is lost with the contractions (Kaczmarski et al. (2022)). The geometric mapping is as follows:

(A) Cylindrical structure in the reference configuration (left) with a vertical central axis and its deformation in the current configuration (right). The deformed configuration is fully parameterized by the centerline r (Z) (resulting from the distortion of the central axis) and the deformation of each cross-section. (B) Schematic representation of the body shape of the C. elegans embryo with the cortical epidermis and the four muscles. The fibers are embedded in the cortex. The blue part representing the epidermis shows the outer distribution of the actin organized into horizontal hoop bundles when the muscles are not activated. The yellow part includes the vertical red muscles, represented by axial fibers. (C) Bending of a Giraffe neck. (D) Torsion of an Elephant trunk. (E) Plant vine twisting. (F) to (H) Deformation configuration under different activations obtained by our simulations for bending and torsion of large rods, twisting and torsion of thin rods.

where the ai represents the deformation in each direction of the section with ai (0, 0, Z) = 0 so that the Z-axis maps to the centerline r (Z). The small quantity ε is the ratio between the radius R and the length L of the cylinder. Referring to the model proposed by B. Kaczmarki et al. (Kaczmarski et al. (2022)), we define the initial configuration ℬ0 with material points (X, Y, Z), and the mapping function χ (X) links the initial configuration ℬ0 to the current configuration ℬ0. The geometric deformation gradient is F = Gradχ = FeG (Nardinocchi and Teresi (2007)), where G represents the active strain generated by the actomyosin or the muscles, and Fe is the elastic strain tensor. Because of the two stages through which the C. elegans elongates, we need to evaluate the influence of the C. elegans actin network during the early elongation before studying the deformation at the late stage. So, the deformation gradient can be decomposed into: F = FeG1G0 (Goriely and Ben Amar (2007)) where G0 refers to the pre-strain of the early period and G1 is the muscle-actomyos in supplementary active strain in the late period. Actin is distributed in a circular pattern in the outer epidermis, see Fig.3(B), so the finite strain G0 is defined as G0 = Diag (1, g0(t), 1), where 0 < g0 < 1 is the time-dependent decreasing eigenvalue, operating in the actin zone and is equal to unity in the other parts: g0(0) = 1. In the case without pre-strain, G0 = I. The deformation gradient follows the description of Eq.(1) :F = Fij di ⊗ ej, i ∈ {1, 2, 3} and j ∈ {R, Θ, Z}.

When considering a filamentary structure with different fiber directions m, these directions are specified by two angles α and β, as outlined in (Holzapfel (2002)): m = sin α sin β eR + sin α cos β e𝚯 + cos α eZ, α, β ∈ [−π/2, π/2], where α and β lie in the range [−π/2, π/2]. For muscle fibers, αm = 0 and βm = 0, whereas for hoop fibers in the actin network, αa = π/2 and βa = 0. When the muscles are activated and bend the embryo, the actin fibers become inclined with a slope α, such that −π/2 < αa < π/2, αa ≠ 0, and βa = 0. Each active strain is represented by a tensor Gi = I + εgimi ⊗ mi, where gi represents the activity (gm for the muscles and ga for acto-myosin), and since both are contractile, their incremental activities are negative. Additional calculation details are provided in Section Methods and Materials.

To relate the deformations to the active forces induced by muscles and the acto-myosin network, we assume that the embryo can be represented by the simplest nonlinear hyperelastic model, called Neo-Hookean, with a strain-energy density given by . In cylindrical coordinates, the total energy of the system and the auxiliary energy density is: where J = det Fe ′ p is a Lagrange multiplier which ensures the incompressibility of the sample, a physical property assumed in living matter. When the cylinder involves several layers with different shear modulus µ and different active strains, the integral over S covers each layer. To minimize the energy over each section for a given active force, we take advantage of the weakness of ε and expand the inner variables ai, pi, and the potential V to obtain the auxiliary strain-energy density. Given that the Euler-Lagrange equations and the boundary conditions are satisfied at each order, we can obtain solutions for the elastic strains at zero order a(0) and at first order a(1). Finally, these solutions which will represent a combination of bending and torsion deformation will last the time of a muscle contraction while the acto-myosin will continue its contractile activity. The evaluation of the elastic energy under acto-myosin-muscle activity at order ε4 can be compared with the equivalent energy of an extensible elastic rod and each basic deformation will be identified (Kirchhoff and Hensel (1883); Mielke (1988); Mora and Müller (2002, 2004); Moulton et al. (2020a)). The typical quantities of interest are the curvature reached during a bending event or a torsion as well as the overall elongation ζ. In the C. elegans embryonic system, these quantities result from the competition between the active strains due to muscles and the myosin motors and the elasticity of the body. Similar active matter can be found in biological systems, from animals to plants as illustrated in Fig.3(C)-(E), they have a structure that generates internal stress/strain when growing or activity. Combining anatomy and measurement techniques, we can transform the mechanics of the body under study into a soft sample submitted to localized internal active stresses or localized internal active strains and then deduce its overall deformation mechanisms, some examples are presented in Fig.3(F)-(H).

The early elongation induced by the acto-myosin

Experimental measurements during the early elongation stage reveal the embryonic diameter and the active or passive stresses, detected by fracture ablation in the body vary with the elongation fold (Vuong-Brender et al. (2017a); Ben Amar et al. (2018)). While previous studies have extensively examined this initial stage, we have chosen to revisit it within the context of our geometry, aiming to attain complete control over our structural modeling, encompassing the geometry, shear modulus, and activity of the acto-myosin prior to muscle contraction. At the outset of the second stage, owing to the early elongation, the embryo experiences a pre-strain, which we describe in our model as G0. The geometry and mechanical information are depicted in Fig.4(A).

(A) The schematic of early elongation and the cross-section of C. elegans. In the cross-section, the black circled part is the actin region (R2′ < R < R3, with shear modulus µa), and the blue part is the epidermis layer (R2 < R < R2′, shear modulus µe). The central or inner part (0 < R < R2, shear modulus µi) except the muscles which are stiffer. (B) Predictions of normalized embryo radius evolution during early elongation by the pre-strain model compared with experimental data from (Vuong-Brender et al. (2017a)). For the model, please refer the Eq.(20) in Appendix 2. (C) Blue dots: extraction of the parameter g0(t) from Eq.(30) and Eq.(32) in the Appendix 2. Blue dash line, refer to the Eq.(3).

Here, we need to first accurately determine G0 by analyzing the experimental data (Vuong-Brender et al. (2017a); Ben Amar et al. (2018)). The cylinder can be divided into three distinct sections: the outer layer is the actin cortex, the thin ring where actin bundles concentrate and work, located within a radius range between R2′ and R3 (as shown in Fig.4(A)); the middle layer (R2 < R < R2′), which is the epidermis but without actin; and the inner part (0 < R < R2), where the muscle is located, along with some internal organs, tissues, and fluids. So, we treat the outer and middle layers as incompressible, but the inner part as a compressible material, except the muscles. The initial deformation gradient: F0 =Diag(r′(R), r(R)/R, λ), and G0 = Diag(1, g0, 1) with 0 < g0 < 1 in the actin layer but with g0 = 1 in the part without actin. G0 represents the circumferential strain exerted by actin during the early elongation and is a slowly varying function of time. r(R) is the radius after early elongation. By applying the principles of radius continuity, radial stress continuity, and incorporating the zero traction condition on the face of the cylinder, we can determine that g0 = 0.88 when the elongation λ = 1.8, i.e., at the completion of the early elongation. As illustrated in Fig.4(B), the results of our model and the experimental data are in good agreement, this demonstrates the consistency of the geometric and elastic modeling together with the choice of a pre-strain represented by G0 which gives a good prediction of the early elongation. More details are displayed in Appendix 3.

From the first stage elongation represented by the blue dots and the blue dashed line in Fig.4(C) we can extract the time evolution of the contractile pre-strain g0(ti) derived from our elastic model. To explain quantitatively g0(t), we suggest a phenomenological dynamical approach for the population of active myosin motors. This equation considers the competition between the recruitment of new myosin proteins from the epidermis cytoskeleton necessary to extend the embryo and the debonding of these myosins from the actin cables, which is damped by the compressive radial stress. It reads: with Xg = 1 − g0(t), p1 is the ratio between the free available myosin population and the attached ones divided by the time of recruitment (given in minutes), while p2 is the inverse of the debonding time of the myosin motors from the cable: p2 = 6 min−1. The debonding time increases (or decreases) when the actin cable is in radial compressive (or tensile) stress, see Appendix 3, Eq.(35). τυ is the visco-elastic time estimated from laser ablation fracture operated in the epidermis (Vuong-Brender et al. (2017b)): tυ = 6s and τp is the time required for the activation of the myosin motors tp = 1200s (Howard and Clark (2002)). This equation is similar to the model derived by Serra et al (Serra et al. (2021)) for the viscous stress occurring in gastrulation.

Notice that only p1 and p3 must be obtained by comparison of g0(t) deduced from Eq.(3) with the values deduced from our elastic model. The result of Eq.(3) with p1 = 0.6 and p3 = 0.75 are shown in Fig.4(C) with a rather good agreement.

Shape of the embryo under muscles and acto-myosin contraction

The regulation of the muscle contraction in C. elegans detected experimentally (Yang (2017)) indicates a cyclic process, where two muscles contract on one side of the embryo quite simultaneously and then stop, while on the opposite side, the two muscles start their contractions, but only after a delay. Let us consider first that only muscles are active (see the schematic Fig.5(A) for the structure, then (B) and (C) for the bending). In this case, due to the geometry, only a bending deformation occurs on the left for active muscles localized on left and then on the right for the symmetric muscles on right.

(A) Schematic diagram of C. elegans muscle fibers and its cross section, and it does not show the actin fibers. Four muscle bands exist in the yellow layer. But, the yellow region is not an actual tissue layer and it is simply to define the position of muscles. (B) Deformation diagram, when left side muscles M1(θ1 ≤ θ0 ≤ θ2) and M2(θ3 ≤ θ0 ≤ θ4) (C) Deformation diagram, when right side muscles M3(θ5 ≤ θ0 ≤ θ6) and M4 (θ7 ≤ θ0 ≤ θ8). (D) Schematic diagram of C. elegans actin fibers and cross-section. (E) Once the muscle is activated, the actin fiber orientation changes from the ‘loop’ to the ‘slope’, which results in torque. (F) Schematic diagram of torsional and bending deformation.

To be more quantitative, we assume that the left side muscles are activated during a short period with an active constant strain value gm in the region M1 and M2, as shown in Fig.5(A)-(B); if the muscles are perfectly vertical, αm = βm = 0 in the initial configuration. In fact, the two muscles on the same side are not always fully in phase and one may present of small delay. For simplification, we assume them are perfectly synchronous. During the full initial period where muscles are not activated, the actin fibers are distributed in a horizontal loop on the outer surface of the epidermis, but once the muscle starts to contract, the acto-myosin network will be re-orientated (Lardennois et al. (2019)). The fibers will be then distributed in a sloping pattern causing eventually the twisting of the embryo, see the schematic diagram shown in Fig.5(D)-(F). When this region is activated with a constant strain value, ga, the angle of the actin fibers will change following the amplitude of the bending by the muscle contraction. In this situation, the angle of the actin fibers may change from but βa is not modified and βa = 0.

Throughout the entire process, the muscle and acto-myosin activities are assumed to work almost simultaneously. Our modeling allows us to evaluate the bending and torsion generated independently by muscles and actin bundles, culminating in a complete deformation under coupling. Furthermore, the angle of the acto-myosin fibers varies during muscle contractions. We maintain a constant activation of the acto-myosin network and gradually increase muscle activation.

As a result, the bending of the model will be increased, causing a change in the angle of the actin fibers, ultimately yielding a deformation map that is presented in Fig.6(A)-(C). Upon increasing myosin activation, we observed a consistent torsional deformation (Fig.6(E)) that agrees with the patterns seen in the video (Fig.6(D)). However, significant torsional deformations are not always present. In fact, other sources can induce torsion as the default of symmetry of the muscle axis, which we will discuss later. In Fig.6(F)-(G), we demonstrate that the curvature provided by the model increases as muscle activation increases and that the torsion is not simply related to the activation amplitude since it also depends on the value of the angle αa, reaching a maximum at approximately π/4. Detailed calculations are shown in Section Methods and Materials and Appendix 4.

Deformed configurations for different activation for muscles, (A) gm = −0.02, αa = π/3, ga = −0.01. (B) gm = −0.05, αa = π/4, ga = −0.01. (C) gm = −0.08, αa = π/6, ga = −0.01. (D) The graphs were captured from the Hymanlab, and the website: https://www.youtube.com/watch?v=M2ApXHhYbaw. The movie was acquired at a temperature of 200C using DIC optics. (E) gm = −0.1, αa = π/4, ga = −0.7. (F) Curvature is plotted as a function of muscle activation. (G) Torsion is plotted as a function of the actin activation and angle of actin fibers.

Energy transformation and Elongation

During the late elongation process, the four internal muscle bands cyclically contract in pairs (Yang (2017); Williams and Waterston (1994)). Each contraction of a pair increases the energy of the system under investigation, which is then rapidly released to the body. This energy exchange causes the torsion-bending energy to convert into elongation energy, leading to a length increase during the relaxation phase, as shown in Fig.1 of the Appendix 5. With all deformations obtained, Eq.(2) can be used to calculate the accumulated energy Wc produced by both the muscles and the acto-myosin activities during a contraction. Subsequently, when the muscles on one side relax, the worm body reverts to its original shape but with a tiny elongation corresponding to the transferred elastic energy. This new state involves the actin network adopting a ‘loop’ configuration with a strain of εga1 once relaxation is complete. If all accumulated energy from the bending-torsion deformation goes towards elongating the worm body, the accumulated energy Wc and energy Wr following muscle relaxation are equivalent. The activation of actin fibers ga1 after muscle relaxation can be calculated and determined by our model.

Once the first stage is well characterized, we can quantify the total energy resulting from both muscles and acto-myosin after each contraction. Then, we used our model to predict elongation in the wild-type C. elegans, unc-112(RNAi) mutant and spc-1(RNAi) pak-1(tm403) mutant and further compared the results with experimental observations. The length of the wild type elongates from approximately 90 µm to 210 µm during the muscle-activated phase (Lardennois et al. (2019)), the phase lasts about 140 minutes, and the average time interval between two contractions is about 40 seconds (Yang (2017)), so the number of contractions can be estimated to be around 210 times. Unfortunately, due to the difficulty to realize quantitative experiments on an embryo always in agitation in the egg shell between bending, torsion, and also rotations around its central axis, one can hypothesize several scenarios: at each step i between state Ai,0 to Ai,1 and then Ai,2, the whole mechanical muscle-myosin energy is transferred to the elongated step Ai,2, leading to a small δζi.

Considering the experimental results shown in Fig.7(B), we determine the optimal values for the activation parameters: gm = −0.15 and ga = −0.01 assuming that all the accumulated energy during muscle activation is transferred to elongation(Wr = Wc). The elementary elongation δζi will be gradually increased with time, which is shown as the black line in Fig.7(A). At the beginning, δζi is about 0.5µm, but at the end of this process, δζi is about 1.5µm, indicating that the worm will elongate up to 290µm. The result is significantly higher than our actual size 210µm. When the elongation proceeds, we assume a transfer of energy between bending-torsion-contraction and elongation but it may be not fully effective, which means a significant part of the energy is lost. From the experimental data, we evaluate that the energy loss gradually increases, from full conversion at the beginning to only 40% of the accumulated energy used for elongation at the end of the process (Wr = 0.4Wc). It induces for δζi a first increase and then a decrease, which is shown as the blue line in Fig.7(A) and is responsible for the slowdown around 200 mins. This option, which may be not the only possible one, leads to the estimated elongation having a good agreement with experimental data (see the blue-dashed curve in Fig.7(B)). Indeed it is possible that the C. elegans elongation requires other transformations which will cost energy. As the embryo gradually elongates, energy dissipation and the biomechanical energy required to reorganize the actin bundles may be two factors that contribute to the increased energy loss that underlies the hypothesis.

(A) The elongation for each contraction varies with time. Black line: all energy converted to the elongation, blue line: partial energy converted to the elongation. The activation: gm = −0.15, ga = −0.01. (B) The model predicted results agree well with the experimental data of wild-type and different mutant C. elegans embryos (Lardennois et al. (2019)). The activation of wild type model (blue dashed line): gm = −0.15, ga = −0.01. The activation of unc-112(RNAi) (brown dashed line): gm = 0, ga = 0. In the pre-stretch failure case (green dashed line), λ will decrease from 1.8.

Moreover, it was reported in (Norman and Moerman (2002)) that the knockdown of unc-112(RNAi), known for impairing muscle contractions, results in the arrest of elongation of embryos at the twofold stage, indicating that muscles have no activation, gm = 0 in our model, and no accumulated energy can be converted into elongation. Another mutation concerning the embryos consisting of mutant cells with pak-1(tm403), known to regulate the activity of myosin motors leads to a retraction of the embryo, so the pre-stretch caused by myosin will not be maintained and will decrease. These aforementioned findings are fully consistent with a variety of experimental observations and are shown in Fig.7(B).

Embryo rotations and dissipation

The main manifestation of the muscle activity, independently of the elongation, is probably the constant rotations of the embryo despite its confinement in the egg. This can be explained by a small angular deviation of the muscle sarcomeres from the central axis due to their attachment to the inner boundary of the cell epidermis, the so-called “dense bodies” (Moerman and Williams (2006)). Since they cross the horizontal plane at approximately ±45◦ and the deviation βm from the anterior-posterior axis is estimated to be about 6◦ each active muscle on the left (or on the right) contributes to the torque via a geometrical factor about ag = sin(6π/180) cos(π/4). Then a simple estimation of the muscle activity in terms of torque reads Λm ∼ µmπR3smpmag(εgm) where sm is the surface of the muscle pair on the left (or right) compared to the section of the cylinder: sm = 0.025 and pm is the distance of the muscles from the central axis of the embryo: pm = 0.75 while gm = 0.15 according to the analysis of the elongation. So the muscles on one side contribute to a torque Λm along the axis of symmetry given by Λm = 4.657µmπR3 · 10−5.

Let us consider now the dissipative torque, assuming that the dynamics of rotation is stopped by friction after one bending event. Two cases can be considered: either the dissipation comes from viscous flow or from the rubbing of the embryo when it folds. The fluid dissipation results from the rotations in the interstitial fluid inside and along the egg shell since the anterior-posterior axis remains parallel to the egg-shell (Bhatnagar et al. (2023)) The interstitial fluid, of viscosity η contains a significant amount of sugar and other molecules which are required for embryo survival and then is more viscous than water (Labouesse (2023)). However, values for sucrose or sorbitol at the concentration of 1 mole/liter indicates a viscosity of order a few times the viscosity of water, which is 1 mPas. For example at 0.9 mole/liter and temperature of 20◦, an aqueous solution with sorbitol has a viscosity of 1.6 mPas, which can be extrapolated to η = 1.9 mPas at 1.2 mole/liter (Jiang et al. (2013)). The estimation given by an embryo located in the middle of the egg-shell, gives a weaker viscous torque once evaluated by according to a classical result reported by Landau et al. (Landau and Lifshitz (1971, 2013)). It has to be mentioned that this estimation assumes that the two cylinders: the egg shell and the embryo have the same axis of symmetry and concerns the beginning of the muscle activity where the radius is about 8.2 µm, the length is 90 µm,the radius of the shell about Regg = 15µm and the length Legg = 54µm, see Fig.6(D). The angular velocity Ωe is more difficult to evaluate but it is about 90◦ per two seconds deduced from videos. When the embryo approaches the egg shell, the friction increases, and two eccentric cylinders of different radii have to be considered with the two axes of symmetry separated by a distance d. The hydrodynamic study in this case is far from being trivial, and seems to have been initiated first by Zhukoski (Zhukoski (1887)) who suggested the use of bipolar coordinates for the mathematical treatment. Many following contributions established with different simplified assumptions have been published after and the study was fully revisited by Ballal and Rivlin (Ballal and Rivlin (1976)). Here we focus on the limit of a small gap δ between the rotating body-shape and the egg and by considering an asymptotic analysis at small δ of the general result derived in (Ballal and Rivlin (1976)). Thus, the viscous torque reads is the zone of contact with the egg and d = Regg − δ − R. This approach is an approximation since the embryo has more of a torus shape than a cylinder (Yang Xinyi et al. (2023)) but the evaluation of the dissipation is satisfactory for δ = 0.5µm, Ωe ∼ π/4 s−1 and µ = 105P a. Coming back to the first model of dissipation with the same data, the ratio between the dissipative viscous torque and the active one gives: Λυ/Λm = 0.02, which is obviously unsatisfactory. Finally, the dissipative energy ? during one bending event leading to an angle of π/2 is ξdiss = 1/2Λm ×(π/2)2 which represents 4% of the muscle elastic energy during the bending so at the beginning of the muscle activity (Appendix 5, Eq.(45)), the dissipation exits but is negligible. At the very end of the process, this ratio becomes 60% but as yet mentioned our estimation for the dissipation becomes very approximate, increases a lot due to the embryo confinement, and does not involve the numerous biochemistry steps necessary to reorganize the active network: acto-myosin and muscles.

Discussion

Since the discovery of the muscle activity before the egg hatch of the C. elegans embryo, it has become critical to explain the role of mechanical forces generated by muscle contraction on the behavioral and functional aspects of the epidermis. We provide a mechanical model in which the C. elegans is simplified as a cylinder, and the muscle bands and actin that drive its elongation are modeled as active structures in a realistic position. We determine the fiber orientation using experimental observations and then calculate the deformation by tensorial analysis involving the strains generated by the active components. Although a special focus is made on late elongation, its quantitative treatment cannot avoid the influence of the first stage of elongation due to the acto-myosin network, which is responsible for a pre-strain of the embryo. In a finite elasticity formalism, the deformations induced by muscles in a second step are coupled to the level of strains of the initial elongation period. For that, we need to revisit the theory of the acto-myosin contraction and previous results (Ciarletta et al. (2009); Vuong-Brender et al. (2017a); Ben Amar et al. (2018)) to unify the full treatment. In particular, a model for the recruitment of active myosin motors under forcing is presented which recovers the experimental results of the first elongation stage.

The elongation process of C. elegans during the late period is much more complex than the early elongation stage which is caused only by actin contraction. During the late elongation, the worm is distorted by the combined action of muscle and acto-myosin, resulting in an energy-accumulating process. Bending deformation is a phenomenon resulting from unilateral muscular contraction, and during the late elongation, significant torsional deformation is observed, indicating that the bending process induces a reorientation of the actin fibers. It is worth noting that the embryo is always rotating in tandem with the muscle activity making difficult any experimental measurements. However, our model can predict that if the muscles are not perfectly vertical, torque exits that causes rotation and eventually torsion. The accumulated energy is then partly turned into energy for the ongoing action of actin, allowing the embryo to elongate when the muscle relaxes. Both sides of the C. elegans muscles contract in a sequential cycle, repeating the energy conversion process, and eventually completing the elongation process. However, the energy exchange between bending and elongation is limited, among other factors, by the viscous dissipation induced by rotation, which is also evaluated in this study. Not investigated in detail here is the necessary re-organization of the active networks (acto-myosin and muscles) due to this tremendous shape transformation of the embryo. In parallel to elongation, the cuticle is built around the body (Page and Johnstone (2007)). This very thin and stiff membrane ensures protection and locomotion post-hatching. Clearly, these processes will perturb muscle activity. These two aspects which intervene in the final stage of the worm confinement play a very important role at the frontier across scales between genetics, biochemistry and mechanics.

Finally, the framework presented here not only provides a theoretical explanation for embryonic elongation in C. elegans, it can also be used to model other biological behaviors, such as plant tropism (Moulton et al. (2020b)) and elephant trunk elongated (Schulz et al. (2022b,a)) and bending. Our ideas could potentially be used in the emerging field of soft robotics, like octopus legs-inspired robots (Kang et al. (2012); Nakajima et al. (2013); Calisti et al. (2015)), which is soft and its deformation induced by muscles activation. We can reliably predict deformation by knowing the position of activation and the magnitude of forces in the model. Furthermore, residual stresses can be incorporated into our model to fulfill design objectives.

Methods and Materials

The model has been presented in a series of articles by A. Goriely and collaborators (Goriely and Ta- bor (2013); Moulton et al. (2013, 2020b); Kaczmarski et al. (2022); Goriely et al. (2022)). As one imagines, it is far from triviality and most of the literature on this subject concerns either full cylinders or cylindrical shells in torsion around the axis of symmetry, the case of bending is more overlooked. However, the geometry of the muscles in C. elegans leads automatically to a bending process that cannot be discarded. We take advantage of these previous works and apply the methodology to our model.

The deformation gradient

The axial extension ζ is obtained by r′ (Z) = ζ d3, where ′ denotes the first derivative with respect to the material coordinate Z. From the director basis, the Darboux curvature vector reads: u = u1d1 + u2d2 + u3d3, this vector gives the evolution of the director basis along the filamentary line as: di′ (Z) = ζ u × di, see Fig.3(A).

The finite strain G0 can map the initial stress-free state ℬ0 to a state ℬ1, which reflects early elongation process. After in the residually stressed ℬ1, we impose an incremental strain field G1 maps the body to ℬ2, which represents late elongation process. So, the deformation gradient can be expressed as: F = FeG1G0 (Goriely and Ben Amar (2007)). The deformation gradient F = Fij di ⊗ ej, i ∈ {1, 2, 3} and j labels the reference coordinates {R, Θ, Z} finally reads: where λ is the axial extension due to the pre-strained.

We define G0 = Diag (1, 1 + εc, 1), since g0 = 0.88 can be determined through the early elongation, we write it in the form of 1 + εc to simplify the calculation, c ≈ −0.6 (ε ≈ 0.2).

Let us consider fibers oriented along the unit vector m = (sin α sin β, sin α cos β, cos α) for the incremental strain G1, where α and β characterise the angles with ez and e and can be found in Fig.3(A). The active filamentary tensor in cylindrical coordinates is then G = G0 (I + εg m ⊗ m), see (Holzapfel (2002)) which reads:

The energy function

The auxiliary strain-energy density can be obtained by expanding the inner variables ai pi and the potential V (see Eq.(2)) as: For each cross-section, the associated Euler-Lagrange equations take the following form: where j = 1, 2, 3, k = 0, 1, in association with boundary conditions at each order k = 0, 1 on the outer radius Ri: We solve these equations order by order and consider incompressibility. At the lowest order, the incompressibility imposes the deformation: a(0) = (r (R), Θ, 0) and the Euler-Lagrange equation gives the Lagrange parameter p(0) = P0(R) which finally read: and the boundary condition σrr(R = 1) = 0 gives the value of P0 = 1/λ. At 𝒪 (ε), the Euler-Lagrange equations are again automatically satisfied, and at 𝒪(ε2), the crucial question is to get the correct expression for and p(1). Based on the previous subsection, the following form is instuited. As before, the Euler-Lagrange equations and incompressibility condition give the two constants and h1, h2, h3 are a function of R. With the solutions for active strain a(0) and a(1), full expressions of them are given in the Appendix 4. The second order energy takes the form: and we can transform it into this form: where Ai, Bi, Ci and Di are functions of R and Θ.

The method for determining deformation

Comparing with the energy of an extensible elastic rod (Moulton et al. (2020a); Kirchhoff and Hensel (1883); Mielke (1988); Mora and Müller (2002, 2004)), we recognize the classic extensional stiffness K0, bending stiffness K1 and K2, torsional stiffness K3 coefficients: where A2, B2, C2 and D2 are related to the shear modulus µ, so for a uniform material with no variations of shear modulus, SK represents the cross-section of the cylinder. But if not, we need to divide different regions to perform the integration.

We now focus on the intrinsic extension and curvatures of the cylindrical object induced by the active strains, this requires the competition of the active forces with the stiffness of the cylindrical bar following the relationships, where Hi calculated by: where A1, B1, C1 and D1 are related to the shear modulus µ,the fiber angles α and β, and the activation g. So, the integration region SH is divided into the different parts of the embryo which all contribute to the deformation.

To calculate the intrinsic Frenet curvature and torsion, we use the following equation: Finally, for the case of activated left side muscles, we can calculate intrinsic extension and curvatures by Eq.16), û1m = 0 and û3m = 0, so the intrinsic Frenet curvature and torsion Eq.(18), and . For the activated actin case, we can calculate intrinsic extension and curvatures by Eq.(16), û1a = 0 and û2a = 0, and the intrinsic Frenet curvature and torsion Eq.(18), and . These quantities have been used to calculate the deformation images, as shown previously in Fig.6.

After obtaining all solutions, we can use these quantities to calculate the accumulated energy Wc in the system after one contraction by Eq.(2). By defining the activation (gm and ga) and the conversion efficiency of energy, we can obtain the activation (ga1) of the actin bundles after once contraction and further calculate the elongation.

All calculation results and parameters are presented in the Appendix.

eLife Assessment

Using the continuum theory of elastic solids, the authors suggest that periodic muscle contraction leads to elongation of C. elegans embryos by storing elastic energy that is subsequently released by extending the embryo's long axis. This important finding could apply to other developmental processes and be exploited in soft robotics. While the presented evidence is in principle convincing, features of the the theory are not explained in sufficient detail.

Reviewer #1 (Public Review):

Summary:

The authors have made a novel and important effort to distinguish and include different sources of active deformations for fitting C elegans embryo development: cyclic muscle contractions and actomyosion circumferential stresses. The combination and synchronisation of both contributions are, according to the model, responsible for different elongation rates, and can induce bending and torsion deformations, which are a priori not expected from purely contractile forces. The model can be applied to other growth processes in initially cylindrical shapes.

Strengths:

The model allows us to fit and deduce specific growth patterns, frequencies, and locations of contractions that yield the observed axial elongation during the 240 min of the studied process.

The deformation gradient is decomposed according to muscle and actomyosin activity, which can be distinguished and quantified. An energy-transferring process allows for the retrieval of the necessary permanent deformations that embryo development requires.

Weaknesses:

Despite the completeness of the model, the explanation of the methodology needs to be improved. Parameters and quantities are not always explained in the main text and are introduced on some occasions in an ordered manner. This makes the comprehension and deduction of methodology difficult. There are some minor comments that are listed below. The most important points are:

-How are the authors sure that there is a torsional deformation? Without tracking the muscle fibers, bending with respect to different angles for different Zs may yield a shape similar to the one in Figure 6E. Furthermore, it is unclear why the model yields torsion deformation. If material points of actomyosin rings do not change in reference configuration, no helicoidal growth should be happening.

-The triple decomposition F=F_e*G_i*G_0 seems to complicate the expressions of growth and requires the use of angles alpha and beta due to the initial deformation G_0. Why not use a simpler decomposition F=F_e*G, where G contains all contributions from actomyosin and muscle contractions in a material frame? This would avoid considering angles alpha and beta.

The section "Energy transformation and Elongation" is unclear. Indeed, stresses need to relax, otherwise, the removal of muscle and actin activity would send the embryo back to its initial state. However, the rationale behind the energy transfer is not explained. Authors seem to impose W_c=W_r, and from this deduce the necessary actin contraction after muscle relaxation. Why should energy be maintained when muscle relaxes? Which mechanism physically imposes this energy transfer? Muscle contraction could indeed induce elongation if traction forces at the opposite side of the contracting muscle relax. In fact, an alternative approach for obtaining stress relaxation and axial elongation would be converting part of the elastic deformation F_e to a permanent deformation F_p.

-Self contact is ignored. This may well be a shape generator and responsible for bending deformations. The convoluted shape of the embryo in the confined space deserves at least commenting on this limitation of the model.

Reviewer #2 (Public Review):

Summary:

During C. elegans development, embryos undergo elongation of their body axis in the absence of cell proliferation or growth. This process relies in an essential way on periodic contractions of two pairs of muscles that extend along the embryo's main axis. How contraction can lead to extension along the same direction is unknown.

To address this question, the authors use a continuum description of a multicomponent elastic solid. The various components are the interior of the animal, the muscles, and the epidermis. The different components form separate compartments and are described as hyperelastic solids with different shear moduli. For simplicity, a cylindrical geometry is adopted. The authors consider first the early elongation phase, which is driven by contraction of the epidermis, and then late elongation, where contraction of the muscles injects elastic energy into the system, which is then released by elongation. The authors get elongation that can be successfully fitted to the elongation dynamics of wild-type worms and two mutant strains.

Strengths:

The work proposes a physical mechanism underlying a puzzling biological phenomenon. The framework developed by the authors could be used to explain phenomena in other organisms and could be exploited in the design of soft robots.

Weaknesses:

1. This reviewer considers that the quality of the writing is poor. Because of this the main result of this work, how elongation is achieved by contraction, remains unclear to me. In the opinion of this reviewer, the work is not accessible to a biologist. This is a real pity because the findings are potentially of great interest to developmental biologists and engineers alike.

2. The authors assume that the embryo is elastic throughout all stages of development. Is this assumption appropriate? In my opinion, the authors need to critically discuss this assumption and provide justification. Would this still be true for the adult? If so could the adult relax back to the state prior to elongation? The embryo should be able to do that, if the contractility of the epidermis were sufficiently reduced, right?

3. The authors impose strains rather than stress. Since they want to understand the final deformation, I find this surprising. Maybe imposing strain or stress is equivalent, but then you should discuss this.

4. Does your mechanism need 4 muscle strands or would 2 be sufficient?

5. It is sometimes hard to understand, whether the authors are talking about the model or the worm.

Author response:

The following is the authors’ response to the original reviews.

Recommendations for the authors:

Reviewer #1 (Recommendations For The Authors):

Line 144, after eq. (1). Vectors d_i need to be defined. Are these the mapping of vectors e_i due to the active deformation? It would be useful to state then that d_3 is aligned with r'.

Thank you for your suggestion, and the definition has been added to lines 146-149 for a better understanding of the model.

Line 144.Authors state a_i(0,0,Z)=0. Shouldn't this be true also for any angle, i.e., a_i(0,Theta, Z)=0?

Thank you, we have revised it in line 144.

Line 156. G_0 is defined as Diag(1,g_0(t), 1), which seems to be using cylindrical coordinates. Previously, in line 147, vector argument X of \chi is defined with Cartesian coordinates (X,Y,Z). Shouldn't these be also cylindrical?

We are very sorry for this error, our initial configuration is defined with cylindrical coordinates, we have revised it in the manuscript line 151.

Line 162. "where alpha and beta lie in the range [-pi/2, pi/2]" has already been indicated.

Thank you for your mention, we have deleted duplicate information in line 166.

Line 171. W is defined as the strain energy density, while in equation (2), symbol W is the total energy (which depends on the previous W). Letters for total elastic and strain energy must be distinguished.

Thank you, we have changed the letter for total energy in Eq.(2).

Line 176. "we take advantage of the weakness of" -> "we take advantage of the small value of".

We have revised it in line 179.

Line 177. Why is there a subscript i in p_i? If these do not correspond to penalty p, but to parameters in eqn (3), the latter should have been introduced before this line.

We have revised this error in line 180.

Line 186. "as the overall elongation \zeta". This parameter, axial extension, has not been defined yet.

Thank you for your mention, the definition of \zeta is now given in line 146.

Figure 4. Why are the values of g_0 from the elastic model and equations (30)-(32) so non-smooth? Clarify what is being fit and what is the input in the latter equations. Final external radius R_3? Final internal radius R_1'?

(1) To mimic the embryo, we consider a multi-layered cylindrical body so that the shear modulus of each layer is different. The continuity of both deformations and stresses is imposed (see Eq.(26)-Eq.(30). This is the usual treatment for complex morpho-elastic systems. Obviously, $g_0$ originates from the actomyosin cortex so it appears only in the corresponding layer. Finally, all physical quantities such as deformations and stresses must be continuous.

(2) The final outer radius is R_3, which represents the outer radius of C. elegans embryos. In addition to R_3, what we need to consider in this model are R_1’=0.7, R_1’=0.768, R_2=0.8 and R_2’=0.96, these definitions have been added in the caption of Appendix 2—figure 1.

Line 663, equation (19). Parameter mu is multiplying penalisation term with p, while in equation (2) mu is only affecting the elastic part.

These two different ways of expressing the energy function will ultimately affect the value of p, but the two p are not the same quantities, so they will not affect our results. To avoid misunderstandings, we will replace p in equation (19) with q.

Reviewer #2 (Recommendations For The Authors):

As mentioned in my public summary, I find the writing really not adequate. I provide here a list of specific points that the authors should in my opinion address. As a general comment, I would delete many instances of 'the'.

First, here are figures and whole paragraphs that do not seem to bring anything to the understanding of the phenomenon of C. elegans elongation, notably, Figs. 2, 3C-H, 5m, and 6. Figures 6G and 7 are the only figures containing results it seems. Some elements of the figures are repeated, for example, the illustration of the system's cross-section in Figs 3 and 5.

Thank you for your suggestion, we have made some adjustments to our images to remove some of the duplicate information.

Second, and this is my most important criticism: the mechanism of elongation by releasing elastic stress introduced by muscle contraction is not explained in clear terms anywhere in the text. At least, I was unable to understand it. On p 10 you write "This energy exchange causes the torsion-bending energy to convert into elongation energy, (...)" How this is done is not explained. I assume that the reference state is somehow changed through muscle contraction. The new reference state probably has a longer axis than the one before, but this would then be a plastic deformation and not purely elastic as claimed by the authors (ll 76: "This work aims to answer this paradox within the framework of finite elasticity without invoking cell plasticity (...)"). Is torsion important for this process or is it 'just' another way to store elastic energy in the system?

We perfectly explain most of the exchange of energy between bending, torsion and elongation: indeed, we quantify all aspects of this transformation as the elastic elongation energy, and the dissipation processes which will cost energy. The dissipation evaluated here concerns the rotation of the worm due to the muscle geometry and the viscous friction at the inner surface of the egg. Torsion seems to appear in the late stages and only in some cases. As we show, it comes from a torque induced by the muscles which are not vertical. vertical. Finally, our quantitative predictions of the modelling which recovers most of the experimental published results.

Third, there are a number of strange phrasings and the notation is not helpful in places.

We feel sorry for that, the manuscript is now more precise.

Fourth, the title promises to explain how cyclic muscle contractions reinforce acto-myosin motors. I can't see this done in this work.

The fact that the acto-myosin is reorganized between two sequences of contraction justifies the title. The complete reorganization of the actomyosin network would require a chemico-mechanical model that is not achieved here, perhaps in future work as data become available.

In addition:

We have chosen to respond globally rather than point by point to the referee’s recommendations.

Typographic errors and vocabulary

All English corrections and typos are now included in the main text.

Figures and captions:

Figures and captions have been improved.

Figure 1: Make the caption and the illustration more coherent. For example, only two cell types are distinguished; in the caption, you mention lateral cells, in the sketch seam cells. What is the difference between acto-myosin and muscle contraction? Muscle contraction is also auto-myosin-based.

(1) The caption for Fig.1 is revised.

(2) From a mechanical point of view, actomyosin bundles in C elegans are orthoradial, whereas muscles are essentially parallel to the main axis of the body are essentially parallel to the main axis of the body, so the geometry is completely different and of extreme importance for deformation. Muscle contractions are quasi-periodic, we do not know the dynamics of the attached molecular motor of myosin. So of course, both contain actin and myosin (not exactly the same proteins), but our model is sensitive to more macroscopic properties.

Figure 2: I do not find this figure helpful. I might expect such a figure in a grant proposal, but much less in an article.

Figure 2 shows the strategy of our work, we hope that readers can see at a glance what kind of analysis has been done through this figure: since our work is divided into several parts, readers can also unravel the logic through this scheme after reading the whole manuscript. So, this diagram is a guide, and it may be helpful and necessary.

Figure 3: Figure 3 A, right: What is the dashed line? B You indicate fibers, but your model does not contain fibers, does it? How do I get from the cube to the deformed object? What is the relation of C-H with the rest of the work? Furthermore, you mention seam cells in Fig. 1, but they are absent here. Why can you neglect them? Why introduce them in the first place? E What is a plant vine? F-H What rods are you referring to? Plants do not have muscles, right?

We have modified this figure, and the original Figure 3 now corresponds to Figures 3 and 4.

(1) The dashed line is the centerline after deformation.

(2) The referee is wrong: our model represents the fibers by a higher shear modulus for the actomyosin cortex and for the muscles (see Table Appendix 1) and G_1 reflects the activities of the muscle and actin fibers.

(3) The cube in Figure 3 is a mathematical 3D volume element that is subjected to stresses. Hyperelasticity modelling is based on such a representation.

(4) C-H(new version: Fig.4 A-F): These images show similar deformations: bending and torsion as our C. elegans study. These figures indicate that such deformations are quite common in nature, even if the underlying mechanism is different.

(5) This is a point we have already mentioned: we ignore the difference between the different types of epidermal cells and average their role in the early and second stages of elongation.

(6) The plant vine is the 'botanical vine', see Goriely's article and book.

(7) F-H(new version: Fig.4 D-F) do not have fixed rods, we set a curvature and torsion to fit the actual biological behavior.

(8) Plants do not have muscles, but they grow, and our formalism for growth, pre-strain and material plasticity is very similar to the hyper-elasticity formalism.

Figure 4: Fig .4 A: "The central or inner part (0 < 𝑅 < 𝑅2, shear modulus 𝜇𝑖) except the muscles which are stiffer." I do not understand.

In the new version, this figure corresponds to Fig.5. The shear modulus of the intrinsic part is very small, but the muscles are harder so we have to consider them separately, we have revised this sentence to avoid misunderstanding.

Figure 5: Fig 5 A and D: The schematic of the cross-section has appeared already in the previous figure. No need to repeat it here. The same holds for the schematic of the cylindrical embryo. Caption: "But, the yellow region is not an actual tissue layer and it is simply to define the position of muscles." Why do you introduce the yellow region at all? I do not think that it clarifies anything. "Deformation diagram, when left side muscles M_1 and M_2." Something seems to be missing here. Similarly in the next sentence. "the actin fiber orientation changes from the 'loop' to the 'slope'" Do the rings break up and form a helix?

In the new version, this figure corresponds to Fig.6.

(1) We have made revisions to these figures.

(2) The yellow part can show the accurate location of four muscles, which is important for our model and further calculations.

(3) We have revised this sentence in the caption of Fig. 6.

(4) Actin rings do not change to a helix pattern, they will be only sloping.

Figure 6: Fig 6 A-C These panels do not go beyond Fig 5B. Fig 6D: what are these images supposed to show? They are not really graphs, but microscopy images. The caption is not helpful to understand, what the reader is supposed to see here. Fig 6F: do you really want to plot a linear curve?

In the new version, Fig.5 and Fig.6 respectively correspond to Fig.6 and Fig.7.

(1) Fig.6 shows the simulated images, and Fig.7 A-C is the real calculation results, they are different.

(2) Fig.7 D can show the real condition during C. elegans late elongation, here, we would like to show the torsion of the C. elegans.

(3) Yes, it is our result.

Discussions concerning the biological referee questions:

Ll 75: “how the muscle contractions couple to the acto-myosin activity" Again I find this misleading because muscle contraction relies on auto-myosin activity. Probably, you can find a better expression to refer to the activity of the actomyosin network in the epidermis. Do you propose any mechanism for how muscle contraction increases epidermal contractility? This does not seem to be the mechanism that you propose for elongation, is it?

The actomyosin activity will not stop because of the muscle contraction. Obviously, these two processes cannot be independent. The energy released by a muscle contraction event can and must contribute to the reorganization of the actomyosin network that occurs during the elongation process. Indeed, despite the fact that the embryo elongates, the density of actin cables appears to be maintained, which automatically requires a redistribution of actin monomers. We propose a scenario in which muscle contraction increases actomyosin contractility via energy conversion. We show that after unilateral contraction there is an energy release for this once all dissipation factors are eliminated. We invite the reviewer to re-examine Figure 2 and invite biologists to seriously evaluate the density of molecular motors attached to the circumferential actin cable throughout the stretch process.

Ll 133: "we decide to simplify the geometrical aspect because of the mechanical complexity" This is hardly a justification. Why is it appropriate?

Yes, we would like to offer the reader the simplest modelling with a limiting technicity and a limited number of unknown parameters.

L 135: "active strains" Why not active stress?

The two are equivalent, the choice is dictated by the simplicity of deriving quantitative results for comparison with experiments.

L 170: "hyperelastic" Please, explain this term.

It is the elasticity of very soft samples subjected to large deformations. For classic references, see the books of Ogden, Holzapfel and Goriely, all of which are mentioned in our paper.

Major criticism

Eq. 3 and Ll 227: "𝑝1 is the ratio between the free available myosin population and the attached ones divided by the time of recruitment" Why is the time of recruitment the same for all motors? "inverse of the debonding time" Is it the same as the unbinding rate? Why use the symbol p_2 for it? What is p_3?

The model proposed to justify the increase in the activity of the actomyosin motors during the first phase is a mean-field model: thus all quantities are averaged: we are not considering the theory of a single molecular motor, but a collection in a dynamic environment, so we do not need stochasticity here. Equation (3) concerns the compressive pre-strain, which by definition is a quantity varying between $0$ and $1$ and $X_g=1-G$. ... The debonding time is not the same as the debonding rate. The term $p_3$ indicates saturation and is derived from the law of mass action. The good agreement with the experimental data is shown in Fig.5 (A) and (B). An equivalent model has been developed by (M. Serra et al.).

Serra M, Serrano Nájera G, Chuai M, et al. A mechanochemical model recapitulates distinct vertebrate gastrulation modes[J]. Science Advances, 2023, 9(49)

Ll 275: "This energy exchange causes the torsion-bending energy to convert into elongation energy, leading to a length increase during the relaxation phase, as shown in Fig.1 of Appendix 5." You have posed the puzzle of how contraction leads to elongation, and now that you resolve the puzzle, you simply say that torsion and bending energy are converted into elongation. How? Usually, if I deform an elastic object, it will return to its original configuration after releasing the external forces. Why is this not the case here?

Furthermore, the central result of your work is presented in an Appendix!?

We agree with the referee that an elastic object will return to its initial configuration by releasing stress, i.e. by giving up its accumulated elastic energy to the environment. But the elastic energy has to go somewhere, such as heat. We do not dare to say that the temperature of the worm increases during the muscle contractions.

In fact, the referee's comment also assumes that full relaxation of the stresses is possible, so the object is not a multi-layered specimen and/or it is not enclosed in a box. Most living species are under stress, usually called residual stress. Our skin is under stress. Our fingerprints result from an elastic instability of the epidermis, occurring on foetal life as our brain circumvolutions or our vili. . So, it is obvious that stresses are maintained in multilayered living systems. Closer to the case of C. elegans, the existence of stresses has been demonstrated by experiments with laser ablation fractures in the first stage. The fact that the fractures open proves the existence of stress: if not, there is no opening and only a straight line.

Ll 379: "Although a special focus is made on late elongation, its quantitative treatment cannot avoid the influence of the first stage of elongation due to the acto-myosin network, which is responsible for a prestrain of the embryo." This statement is made repeatedly through the manuscript, but I do not understand, why you could not use an initial state without pre-strain.

This is the basic concept of hyperelasticity. The reference state must be free of stress, so we cannot evaluate the first muscle contraction without treating the first elongation stage.

Grammar, vocabulary and writing errors

ll 31: "the influence of mechanical stresses (...) becomes more complex to be identified and quantified" Is the influence of mechanical stress too complex or too difficult to be identified/quantified?

We have revised it in line 31, “The superposition of mechanical stresses, cellular processes (e.g., division, migration), and tissue organization is often too complex to identify and quantify.”

Ll 41: "The embryonic elongation of C. elegans represents an attractive model of matter reorganization without a mass increase before hatching." Maybe "Embryonic elongation of C. elegans before hatching represents an attractive model of matter reorganization in the absence of growth.".

We have revised it in line 41.

L 42: "It happens after the ventral enclosure (...)" Maybe "It happens after ventral enclosure (...)".

We have revised it in line 42.

Ll 52: "The transition is well defined since the muscle participation makes the embryo rather motile impeding any physical experiments such as laser ablation (...)" Ablation of what?

We have revised it in line 53:The transition is well defined, because the muscle involvement makes the embryo rather motile, and any physical experiments such as laser fracture ablation of the epidermis, which could be performed and achieved in the first period (\cite{vuong2017interplay}), become difficult,.

Ll 59: "a hollow cylinder composed of four parts (seam and dorso-ventral cells)" It is not clear, what the four parts are - in the parenthesis, two are mentioned.

We have revised it in line 59. Fig.1 shows the whole structure, dorsal, ventral and seam cells form four parts of the epidermis.

L 78: "several important issues at this stage remain unsettled" At which stage?

It means the late elongation stage, we have added this information in line 78.

Ll 85: "but how it works at small scales remains a challenge." Maybe "but how it works at small scales remains to be understood.".

We have revised it in line 86.

Ll 99: "the osmolarity of the interstitial fluid" The comes out of the blue. Before you only talked about mechanics, why now osmolarity? Also, the interstitial fluid is only mentioned now. It is important for the dissipative effects that you discuss later, right? If yes, then you should probably introduce it earlier.

For a better understanding, we have change osmolarity into viscosity in line 99.

l 120: "The cortex is composed of three distinct cells" Maybe "distinct cell types".

Thank you, and we have revised it in line 120.

L 121: "cytoskeleton organization and actin network configurations" What is the difference between cytoskeleton organization and actin network configuration? Also, either both should be plural or both singular, I guess.

(1) Cytoskeleton (which involves microtubules) forms the epidermis of C. elegans embryos, and the actin network surrounds the epidermis.

(2) Thank you for your suggestion, we have revised it in line 121.

L 130: "which will be introduced hereafter" Maybe "which will be used hereafter".

We have revised it in line 130.

Ll 148: "The geometric deformation gradient" You usually denote vectors in bold face, so \chi should be bold, right? Define d_i in Eq.(1).

Yes, we have added this information in line 147.

L 172: "auxiliary energy density" Please, explain this term.

We have changed "auxiliary energy density" into "associated energy density" in line 175. Energy density is the amount of energy stored in a given system or region of space per unit volume, the associated energy density in our manuscript can help us to do some calculations.

Ll 188: "Similar active matter can be found in biological systems, from animals to plants as illustrated in Fig.3(C)-(E), they have a structure that generates internal stress/strain when growing or activity. (...)" Why such a general statement during the presentation of the results? The second part of the sentence seems to be incomplete.

Answers: We would like to show our method is general, and can be used in many situations. We have revised the wrong sentence in line 192.

Ll 243: "a bending deformation occurs on the left for active muscles localized on left" Maybe "bending to the left occurs if muscles on the left are activated".

Thank you, we have revised it in line 247.

L 250: "we assume them are perfectly synchronous" Maybe "we assume them to contract simultaneously". We have revised it in line 252.

L 258: "the muscle and acto-myosin activities are assumed to work almost simultaneously." Before it was simultaneously, now only almost!? What does almost mean?

Sorry, we would like to express the same meaning in theses two sentences, we have deleted the word ‘almost’ in line 261.

Ll 294: "one can hypothesize several scenarios" After that, only one scenario is described it seems.

Thank you, we have revised this sentence in line 299.

L 341: "and then is more viscous than water" Maybe "and that is more viscous than water".

We have revised it in line 345.

L 373: "before the egg hatch" Maybe "before the embryo (or larva) hatches"?

We have revised the sentence in line 367.

L 409: "elephant trunk elongated" maybe "elephant trunk elongation".

We have revised it in line 412.

Ll 417: "As one imagines, it is far from triviality (...)" Does this remake help in any way to understand better C. elegans elongation? Also maybe "it is far from trivial".

We have revised it in line 423.

Ll 428: "can map the initial stress-free state B_0 to a state B_1, which reflects early elongation process" Maybe: "maps the initial stress-free state B_0 to a state B_1, which describes early elongation".

We have revised it in line 428.

L 429: "After in the residually stressed (...)" Maybe "Subsequently, we impose an incremental strain filed G_1 that maps the state B_1 to the state B_2, which represents late elongation".

We have revised it in line 429.

l 763: "Modelling details of without pre-strain case" Maybe "Case without pre-strain" or "Modelling in the absence of pre-strain" Similarly for l 784.

We have revised them in line 763 and line 784.

Some questions of definition and understanding

Ll 71: "We can imagine that once the muscle is activated on one side, it can only contract, and then the contraction forces will be transmitted to the epidermis on this side." I do not understand the sentence. Muscle activation leads to contraction, there is nothing to imagine here. Maybe you hypothesize that the muscles are attached to the epidermis such that muscle contraction leads to epidermis deformation?

Yes, four muscle bands are attached to the epidermis, as shown in Fig.1. The deformation does not concern only the epidermis but the whole embryo during the bending events. We have modified the sentence to avoid misunderstanding, the sentence change to “Once the muscle is activated on one side, it can only contract, and then the contraction forces will be transmitted to the epidermis on this side.” in line 71.

Ll 110: "However, it is less widely known that its internal striated muscles share similarities with skeletal muscles found in vertebrates in terms of both function and structure" Is it important for what you report, whether this fact is widely known?

Yes, it is our opinion.

Ll 112: "the role of the four axial muscles (...) is nearly contra-intuitive" Is it or is it not? If yes, why?

Yes it is. Muscles exert contractions, so compressive deformations. Their localization are along the axis of symmetry (up to a small deviation) so they cannot mechanically realize the expected elongation, contrary to the orthoradial actomyosin network.

However, elongation of the C. elegans is observed experimentally, so yes, we think the result contraintuitive.

L 116: "fully heterogeneous cylinder" What is this?

It means that the C. elegans embryo does not have the same elastic properties in different parts (or layers).

L 129: "will collaborate to facilitate further elongation" To facilitate or to drive? If the former, what drives elongation?

Contraction of muscles and actin bundles together drive elongation

Ll 141: "the deformation in each section can be quantified since the circular geometry is lost with the contractions" The deformation could also be quantified if the sections remained circular, right?

Yes. However, circularity is lost during each bending event.

Ll 151: "we need to evaluate the influence of the C. elegans actin network during the early elongation before studying the deformation at the late stage. So, the deformation gradient can be decomposed into: (...) where (...) is the muscle-actomyosin supplementary active strain in the late period" I thought you were now studying the early stage?

In this part, we are outlining how we can study the whole elongation (early and late), not just the early elongation stage. To evaluate the deformation induced by the first contraction of the muscles, we need to know the state of stress of the worm prior to this event, so we also need to recover the early period using the same formalism for the same structure.

L 160: "When considering a filamentary structure with different fiber directions" Which filamentary structure are you talking about?

Fig.3 B shows this model and the filamentary structure, which contains the actin and muscle fibers.

Ll 174: "When the cylinder involves several layers with different shear modulus 𝜇 and different active strains, the integral over 𝑆 covers each layer" I do not understand this sentence. Also, you should probably write 'moduli' instead of modulus.

This implies that when integrating over the whole cross-section S, we need to take into account each layer independently with its own shear modulus and sum the results.

L 176: "weakness of 𝜀" Do you mean \epsilon << 1?

Yes

Ll 178: "Given that the Euler-Lagrange equations and the boundary conditions are satisfied at each order, we can obtain solutions for the elastic strains at zero order 𝐚(𝟎) and at first order 𝐚(𝟏)." Are you thinking about different orders in an \epsilon expansion or the early and the late stages of elongation?

Answers: Different orders are considered only for the late elongation study, the early elongation is treated exactly so do not need a correction in \epsilon.

L 197: "fracture ablation" Please, define.

This is an experiment in which a laser is used to make a cut in a small-scale object of study and then the internal stresses are obtained based on the morphology of the cut, please see the Ref ‘Assessing the contribution of active and passive stresses in C. elegans elongation’. We have added this definition in line 200.

Ll 203: What motivated your choice of notations for the radii R_2'? The inner part of the cylinder is fluid? But above you wrote about a solid cylinder. Why should the inner part be compressible?

(1) We need to define the location of actin cables, which concentrate at the outer periphery.

(2) Our model is a hollow cylinder, and the inner part of the cylinder contains internal organs, tissues, fluids, and so on, so we consider it to be a compressible extremely soft material (Line 213).

Ll 212: "𝑟(𝑅) is the radius after early elongation." And during?

R is variable, r(R) depends on R but also on time t, it represents the radius of C. elegans embryos after the onset of elongation, i.e., after acto-myosin and muscle activities begin.

L 232: \tau_p is probably t_p?

Yes.

L 240: "quite simultaneously" Please, be precise.

In practice, it is difficult to define the concept of simultaneous occurrence unless there is rigorous experimental data to show it, but all we can get in the Ref ‘Remodelage des jonctions sous stress mécanique’, is that it occurs almost simultaneously, which we define as quite simultaneously.

Ll 246: "a short period" What does short mean? Why is it relevant?

From the experimental observations and data, we know that each contraction occurs very rapidly: a few seconds so we define a short period for one contraction.

L 263: "the bending of the model will be increased" Is it really the model that is bent?

Yes, the bending deformation predicted by the model, we have revised in line 266.

Ll 265: "we observed a consistent torsional deformation (Fig.6(E)) that agrees with the patterns seen in the video" In which sense do these configurations agree? I do not see any similarity between panels D and E.

Both show a torsion deformation.

L 267: "torsion as the default of symmetry of the muscle axis" I do not understand.

We discuss two cases in this research, one where the muscle follows the axis of the C. elegans in the initial configuration, and the other where the muscle has a slight angle of deflection, and we have added more information in the manuscript (line 270).

Ll 274: "Each contraction of a pair increases the energy of the system under investigation, which is then rapidly released to the body." Do you mean the elastic energy stored in the epidermis and central part of the embryo?

Yes, the whole body.

Ll 284: "The activation of actin fibers 𝑔𝑎1 after muscle relaxation can be calculated and determined by our model." Have you done it?

Yes, we can obtain the value of g_a1, and then calculate the elongation.

Ll 286 I do not understand, why you write about mutants at this place. Am I supposed to have already understood the basic mechanism of elongation? Why do you now write about the first stage?

I would like to show our formalism can model wild-type and mutant C.elegans, and the comparison results are good.

L 302: "The result is significantly higher than our actual size 210𝜇𝑚." How was significance assessed? Your actual size is probably more than 210µm.

Here, we have considered two situations, one is that the accumulated energy is totally applied to the elongation so that the length will be much larger than the experimental result of 210 µm, the length value that we have obtained by calculation. In the other case, we have considered the energy dissipation, which leads to 210 µm.

L 433: "where 𝜆 is the axial extension due to the pre-strained" Maybe ""where 𝜆 is the axial extension due to the pre-stress".

In our manuscript, we define the pre-strain, not the pre-stress.

L 438: "active filamentary tensor" Please, define.

Active filamentary tensor defines the tensor representing the activities of a cylindrical model composed of different orientations fibers.

Version 2

Abstract

Abstract

The paramount importance of mechanical forces in morphogenesis and embryogenesis is widely recognized, but understanding the mechanism at the cellular and molecular level remains challenging. Because of its simple internal organization, Caenorhabditis elegans is a rewarding system of study. As demonstrated experimentally, after an initial period of steady elongation driven by the actomyosin network, muscle contractions operate a quasi-periodic sequence of bending, rotation and torsion, that leads to the final 4-fold size of the embryo before hatching. How actomyosin and muscles contribute to embryonic elongation is investigated here theoretically. A filamentary elastic model that converts stimuli generated by biochemical signals in the tissue into driving forces, explains embryonic deformation under actin bundles and muscle activity, and dictates mechanisms of late elongation based on the effects of energy conversion and dissipation. We quantify this dynamic transformation by stretches applied to a cylindrical structure that mimics the body shape in finite elasticity, obtaining good agreement and understanding for both wild-type and mutant embryos at all stages.

Introduction

Mechanical stresses play a critical role during embryogenesis, affecting complex biological tissues such as the skin, the brain, and the interior of organs. There are many studies in recent decades, for example, the pattern of the intestine (Coulombre and Coulombre (1958); Hannezo et al. (2011); Li et al. (2011); Ben Amar and Jia (2013); Shyer et al. (2013)), the brain cortex (Toro and Burnod (2005); Goriely et al. (2015); Ben Amar and Bordner (2017); Tallinen et al. (2016)), and the circumvolutions of the fingerprints (Kücken and Newell (2005); Ciarletta and Ben Amar (2012)), which have been interpreted as the result of compressive stresses generated by the growth that occurs a few months after fertilization in humans. The superposition of mechanical stresses, cellular processes (e.g., division, migration), and tissue organization is often too complex to identify and quantify. For example, independent of growth, stresses can be generated by different molecular motors, among which myosin II, linked to actin filaments, is the most abundant during epithelial cell morphogenesis (Vicente-Manzanares et al. (2009)) and cell motility (Cowan and Hyman (2007); Olson and Sahai (2009); Palumbo et al. (2022)). Spatial distribution and dynamics of myosin II strongly influence the morphogenetic process (Levayer and Lecuit (2012); Lv et al. (2022)), as demonstrated for Drosophila (Bertet et al. (2004); Blankenship et al. (2006); Saxena et al. (2014); Shindo and Wallingford (2014)) and also for C. elegans embryos (Priess and Hirsh (1986); Gally et al. (2009); Ben Amar et al. (2018)).

Embryonic elongation of C. elegans before hatching provides an attractive model of matter reorganization in the absence of growth. It occurs after ventral enclosure and takes about 240 minutes to transform the lima bean-shaped embryo into the final elongated worm form: the embryo elongates four times along the anterior/posterior axis by (McKeown et al. (1998)). The short lifespan of the egg before hatching and its transparency make this system ideal for studying the forces that exist at or near the cortical epithelium. However, in contrast to Drosophila and zebrafish embryonic development, there is no cell migration, cell division, or significant change in embryonic volume (Sulston et al. (1983); Priess and Hirsh (1986)), only remarkable epidermal elongation drives the entire morphogenetic process of C. elegans in the post-enclosure period. There are two experimentally identified driving forces for the elongation: the actomyosin contractility in the seam cells of the epidermis, which appears to persist throughout the whole elongation process, and the muscle activity beneath the epidermis, which begins after the 1.7 ∼ 1.8-fold stage. The transition is well defined, because the muscle involvement makes the embryo rather motile, and any physical experiments such as laser fracture ablation of the epidermis, which could be performed and achieved in the first period (Vuong-Brender et al. (2017a)), become difficult. Consequently, the elongation process of C. elegans could be divided into two stages: the early elongation and the late elongation, depending on the muscle activation, and Fig.1(A) shows the whole elongation process (Vuong-Brender et al. (2017a)). Previously, the role of the actomyosin network in the seam cells during the C. elegans early elongation was investigated (Ben Amar et al. (2018)). Based on the geometry of a hollow cylinder composed of dorsal, ventral and seam cells, a model including the pre-stress responsible for the enclosure, the active compressive ortho-radial stress combined with the passive stress quantitatively predicts the elongation, but only up to ∼ 70% of the initial length.

(A) Overview of C. elegans embryonic development. Three epidermal cell types are found around the circumference: dorsal, ventral and seam cells. (B) Schemes showing a C. elegans crossn section of the embryo. Small yellow arrows in the left image indicate the contraction force that occurred in the seam cell. Four muscle bands under the epidermis and actin bundles surround the outer epidermis.

For the late elongation phase, the actomyosin network and the muscles together drive the full elongation during this phase (Vuong-Brender et al. (2017a)). The muscles play an important role and mutants with muscle defects are unable to complete the elongation process, even though the actomyosin network functions normally (Lardennois et al. (2019)). Fig.1(B) shows a schematic image of the C. elegans body (Zhang et al. (2011)) with four rows of muscles, two of which are under the dorsal epidermis and the other two are under the ventral epidermis. As observed in vivo, C. elegans exhibits systematic rotations accompanying each contraction (Yang (2017); Yang Xinyi et al. (2023)), and deformations such as bending and twisting. Eventually, the C. elegans embryo will undergo an elongation of 1.8-fold to 4-fold. As the activated muscle contracts on one side, the contractile forces are transmitted to the epidermis on that side. Since striated muscles can only perform cycles of contraction and relaxation, their action will tend to reduce the length of the embryo. Therefore, it is necessary to understand how the embryo elongates during each contraction and how the muscle contractions are coupled to the actomyosin activity. This work aims to answer this paradox within the framework of finite elasticity without invoking cellular plasticity and stochasticity, which cannot be considered as driving forces. In addition, several important questions remain unanswered at the late stage of elongation. First of all, we may also observe a torsion of the embryo, but if the muscle activity suggests a bending, this alone cannot explain alone a substantial torsion. Second, a small deviation of the muscle axis (Moerman and Williams (2006)) is responsible for a number of rotations, how to relate these rotations to the muscle activation (Yang (2017)). Since any measurement on a motile embryo at this scale is difficult, it is useful to explore the mechanism of late elongation theoretically. Furthermore, muscle contraction is crucial for both biological development and activities and has been extensively studied by researchers (Tan and De Vita (2015); De Vita et al. (2017)), but how it works at small scales remains to be understood.

Using a finite elasticity model and assuming that the embryonic body shape is cylindrical, we can evaluate the geometric bending deformation and the energy released during each muscle contraction on one side since after each contraction, the muscles relax and then the muscles on the opposite side undergo a new contraction. This cyclic process results in a tiny elongation of the cylinder along its axis of symmetry with each contraction. Each of these contractions correlates with the rotational movements of the embryo (Yang (2017)). By repeating these pairs of contractions more than two hundred times, a cumulative extension is achieved, but it must be reduced by friction mechanisms, also evaluated by the model. Furthermore, the mechanical model explains the existence of a torque acting on the position of the head or tail by the coupling of the muscle contraction with the orthoradial actomyosin forces. Finally, the small deviation between the muscles and the central axis experimentally detected (Moerman and Williams (2006)) induces cyclic rotations and possibly torsions leading to fluid viscous flow inside the egg. The quantification of all these processes allows the evaluation of physical quantities of the embryo, such as the shear modulus of each component, the interstitial fluid viscosity and the active forces exerted by the actomyosin network and the muscles, which are poorly y known in the embryonic stages.

Our mechanical model accounts for the dynamic deformations induced by the internal stimuli of a layered soft cylinder, allowing us to make accurate quantitative predictions about the active networks of the embryo. Furthermore, our model accounts for the dissipation that occurs in the late period just before the embryo hatches. Our results are consistent with observations of actomyosin and muscle activity (Vuong-Brender et al. (2017a); Ben Amar et al. (2018); Lardennois et al. (2019)). The architecture of our work is illustrated in Fig.2.

Architecture of the program. The program reflects the framework of the research. On the one hand, the proposed model explains the early and late elongation of the C. elegans, on the other hand, the early myosin activity is estimated, the deformations (bending, twisting, rotation) occurring in the late period are recovered, and the estimation of the energy dissipated during muscle activity is achieved.

Results

It is widely recognized that C. elegans is a well-established model organism in the field of developmental biology. However, it is less well known that its internal striated muscles share similarities with vertebrate skeletal muscles in terms of both function and structure (Lesanpezeshki et al. (2021)). Since they are contractile, the role of the four axial muscles (as shown in Fig.1) in the final shaping of the embryo is almost counterintuitive. In this section, we aim to elucidate the action of these muscles when coupled to the actomyosin network, using a purely mechanical approach. To achieve a quantitative understanding, our method requires a detailed description of the deformation geometry of the embryo, where the body shape is represented by a fully heterogeneous cylinder.

Geometric and strain deformations of the embryo

The early elongation of the C. elegans embryo has been previously studied, which is characterized by an inner cylinder surrounded by epithelial cells located in the cortical position (Ben Amar et al. (2018)). The cortex is composed of three distinct cell types - the seam, dorsal, and ventral cells - that exhibit a unique cytoskeletal organization and actin network configuration. Among these, only the seam cells possess active myosin motors that function in the ortho-radial direction, allowing for the circumferential contraction and thereby triggering the early elongation process. In this study, we simplify the cylindrical geometry of the body and treat the epidermis as a unified whole with effective activity localized along the circumference, supported by four muscles distributed beneath the epidermis. The muscles do not play a role in the early elongation phase and have therefore not been considered previously (Vuong-Brender et al. (2017a); Ben Amar et al. (2018)). After this period of approximately a 1.8-fold increase in length, muscles parallel to the main axis and actin bundles organized along the circumference will cooperate to drive further elongation. Table. 1 provides the size parameters of C. elegans that will be used hereafter.

We focus on the total deformation of a full cylinder or a thin rod with a length L greater than the radius R and a central vertical axis along the Z direction. Contrary to previous works (Ciarletta et al. (2009); Ben Amar et al. (2018)), here we decide to simplify the geometrical aspect due to the mechanical complexity. The biological activity induced by the actomyosin network and muscles is represented by active strains, and the global shape is the result of the coupling between elastic and active strains, modulated by dissipation. Active strains are generated by non-mechanical processes (e.g., biochemical processes): myosin motors and striated muscles derive their energy from ATP hydrolysis, which is converted into mechanical fiber contractions. Due to the significant deformation observed, the central line is distorted and becomes a curve in three-dimensional space, represented by a vector r (Z), as depicted in Fig.3(A). Along this curve, perpendicular planar sections of the embryo can be defined, and the deformation in each section can be quantified since the circular geometry is lost with the contractions (Kaczmarski et al. (2022)). The geometric mapping is as follows: where the ai represents the deformation in each direction of the section with ai (0, Θ, Z) = 0 so that the Z-axis is mapped to the centerline r (Z). The small quantity ε is the ratio between the radius R and the length L of the cylinder. The axial extension ζ is given by r′ (Z) = ζ d3, where′ denotes the first derivative with respect to the material coordinate Z. From the director basis, the Darboux curvature vector reads: u = u1d1 +u2d2 +u3d3, this vector gives the evolution of the director basis along the filamentary line as: di′ (Z) = ζ u × di, see Fig.3(A). Based on the model proposed by B. Kaczmarki et al. (Kaczmarski et al. (2022)), we define the initial configuration ℬ0 with material points (R, Θ, Z), and the mapping function χ (X) connects the initial configuration ℬ0 to the current configuration ℬ. The geometric deformation gradient is F = Gradχ = FeG (Nardinocchi and Teresi (2007)), where G is the active strain generated by the actomyosin or the muscles, and Fe is the elastic strain tensor.

(A) Cylindrical structure in the reference configuration (left) with a vertical centerline and its deformation in the current configuration (right). The deformed configuration is fully parameterized by the centerline r (Z) (resulting from the distortion of the central axis) and the deformation of each cross section. (B) Schematic representation of the body shape of the C. elegans embryo with the cortical epidermis and the four muscles. The fibers are embedded in the cortex. The blue part representing the epidermis shows the outer distribution of the actin organized in horizontal hoop bundles when the muscles are not activated. The yellow part includes the vertical red muscles, represented by axial fibers.

Because of the two stages through which the C. elegans elongates, we need to evaluate the influence of the C. elegans actin network during the early elongation before studying the deformation at the late stage. Thus, the deformation gradient can be decomposed into: F = FeG1G0 (Goriely and Ben Amar (2007)) where G0 refers to the pre-strain of the early period and G1 is the muscle supplementary active strain in the late period. Actin is distributed in a circular pattern in the outer epidermis, see Fig.3(B), so the finite strain G0 is defined as G0 = Diag (1, g0(t), 1), where 0 < g0 < 1 is the time-dependent decreasing eigenvalue, operating in the actin zone and is equal to unity in the other parts: g0(0) = 1. In the case without pre-strain, G0 = I. The deformation gradient follows the description of Eq.(1) :F = Fij di ⊗ ej, where i ∈ {1, 2, 3} and j ∈ {R, Θ, Z}.

Considering a filamentary structure Fig.3(B) with different fiber directions m, these directions are specified by two angles α and β, as outlined in (Holzapfel (2000)): m = sin α sin β eR +sin α cos β eΘ + cos α eZ, α, β ∈ [−π/2, π/2]. For muscle fibers, αm = 0 and βm = 0, while for hoop fibers in the actin network, αa = π/2 and βa = 0. When the muscles are activated and bend the embryo, the actin fibers are tilted with a tilt α, such that −π/2 < αa < π/2, αa ≠ 0, and βa = 0. Each active strain is represented by a tensor Gi = I + εgimi ⊗ mi, where gi represents the activity (gm for muscles and ga for actomyosin), and since both are contractile, their incremental activities are negative. Additional calculation details are provided in Section Methods and Materials.

To relate the deformations to the active forces induced by the muscles and the actomyosin network, we assume that the embryo can be represented by the simplest nonlinear hyperelastic model, called Neo-Hookean, with a strain energy density given by . In cylindrical coordinates, the total energy of the system and the associated energy density is: where J = det Fe, p is a Lagrange multiplier that ensures the incompressibility of the sample, a physical property assumed in living matter. If the cylinder contains several layers with different shear modulus μ and different active strains, the integral over S covers each layer. To minimize the energy over each section for a given active force, we take advantage of the small value of ε and expand the inner variables a, p, and the potential V to obtain the associated strain-energy density. Since the Euler-Lagrange equations and the boundary conditions are satisfied at each order, we can obtain solutions for the elastic strains at zero order a(0) and at first order a(1). Finally, these solutions which will represent a combination of bending and torsional deformation, will last for the duration of a muscle contraction as the actomyosin continues its contractile activity. The evaluation of the elastic energy under actomyosin muscle activity of order ε4 can be compared with the equivalent energy of an extensible elastic rod and any fundamental deformation will be identified (Kirchhoff and Hensel (1883); Mielke (1988); Mora and Müller (2002, 2004); Moulton et al. (2020a)). The typical quantities of interest are the curvature achieved during a bending event or a torsion as well as the total elongation ζ. In the C. elegans embryonic system, these quantities result from the competition between the active strains due to muscles and the myosin motors and the elasticity of the body. Similar active matter can be found in biological systems, from animals to plants, as illustrated in Fig.4(A)-(C), they have a structure that generates internal stress/strain for the growth or daily activities. By combining anatomy and measurement techniques, we can transform the mechanics of the body under study into a soft sample subjected to localized internal active stresses or localized internal active strains and then deduce its overall deformation mechanisms, some examples are presented in Fig.4(D)-(F).

(A) Bending of a giraffe’s neck. (B) Torsion of an elephant’s trunk. (B) Twisting of a plant vine. (D) to (F) Deformation configuration under different activations obtained by our simulations for bending and torsion of large rods, twisting and torsion of thin rods.

The early elongation induced by the actomyosin

Experimental measurements during the early elongation stage reveal the embryonic diameter and the active or passive stresses, estimated by the opening of fractures realized in the body by laser ablation (Vuong-Brender et al. (2017a); Ben Amar et al. (2018)). These quantities vary with the elongation. While previous studies have extensively investigated this initial stage, we have chosen to revisit it in the context of our geometry, with the goal of achieving complete control over our structural modeling, which includes the geometry, shear modulus, and activity of the actomyosin prior to muscle contraction. At the beginning of the second stage, the embryo experiences a pre-strain due to the early elongation, which we describe in our model as G0. The geometry and mechanical information are depicted in Fig.5(A).

(A) Schematic of early elongation and the cross section of C. elegans. In the cross-section, the black circular part is the actin region (R2′ < R < R3′ with shear modulus μa), the blue part is the epidermis layer (R2 < R < R2′ with shear modulus μe). The central or inner part (0 < R < R2 in white has shear modulus μi, except for the four muscles, the shear modulus of the four muscles μm is much larger than the inner part. (B) Predictions of normalized embryo radius evolution during early elongation by the pre-strain model compared to experimental data from (Vuong-Brender et al. (2017a)). For the model, see Eq.(20) in Appendix 2. (C) Blue dots: extraction of the parameter g0(t) from Eq.(30) and Eq.(32) in Appendix 2. Blue dash line, see Eq.(3).

Here, we must first accurately determine G0 by analyzing the experimental data (Vuong-Brender et al. (2017a); Ben Amar et al. (2018)). The cylinder can be divided into three distinct sections: the outer layer is the actin cortex, the thin ring where actin bundles concentrate and work, located within a radius range between R2′ and R3 (as shown in Fig.5(A)); the middle layer (R2 < R < R2′), which is the epidermis but without actin; and the inner part (0 < R < R2), where the muscle is located, along with some internal organs, tissues, and fluids. So, we treat the outer and middle layers as incompressible, but the inner part as a compressible material, except for the muscles. The initial deformation gradient: F0 =Diag(r′(R), r(R)/R, λ), and G0 = Diag(1, g0, 1) with 0 < g0 < 1 in the actin layer, but with g0 = 1 in the part without actin. G0 represents the circumferential strain exerted by actin during the early elongation and is a slowly varying function of time. By applying the principles of radius continuity, radial stress continuity, and incorporating the zero traction condition on the face of the cylinder, we can determine that g0 = 0.88 when the elongation λ = 1.8, at the end of the early elongation. As illustrated in Fig.5(B), the results of our model and the experimental data are in good agreement, demonstrating the consistency of the geometric and elastic modeling together with the choice of a pre-strain represented by G0 which gives a good prediction of the early elongation. More details can be found in the appendix 3.

From the first stage elongation represented by the blue dots and the blue dashed line in Fig.5(C) we can extract the time evolution of the contractile pre-strain g0(ti) derived from our elastic model. To explain g0(t) quantitatively, we propose a phenomenological dynamical approach for the population of active myosin motors. This equation takes into account the competition between the recruitment of new myosin proteins from the epidermal cytoskeleton, which is necessary to elongate the embryo, and the detachment of these myosins from the actin cables, which is damped by the compressive radial stress. It reads: where Xg = 1 − g0(t), p1 is the ratio of the freely available myosin population to the attached ones divided by the time of recruitment (given in minutes), while p2 is the inverse of the debonding time of the myosin motors from the cable: p2 = 6 min−1. The debonding time increases (or decreases) when the actin cable is under radial compressive (or tensile) stress, see Appendix 3, Eq.(35). τν is the viscoelastic time estimated from laser ablation fracture operated in the epidermis (Vuong-Brender et al. (2017b)): tν = 6S and τp is the time required to activate the myosin motors tp = 1200S (Howard and Clark (2002)). This equation is similar to the model derived by Serra et al (Serra et al. (2021)) for the viscous stress that occurs during gastrulation.

Note that only p1 and p3 need be obtained by comparing g0(t) deduced from Eq.(3) with the values derived from our elastic model. The result of Eq.(3) with p1 = 0.6 and p3 = 0.75 are shown in Fig.5(C) with a rather good agreement.

Shape of the embryo under muscles and actomyosin contraction

The experimental regulation of muscle contraction in C. elegans (Yang (2017)) suggests a cyclic process in which two muscles on one side of the embryo contract quite at the same time and then stop, while on the opposite side the two muscles begin to contract, but with a delay. Let us first consider that only the muscles are active (see the schematic Fig.6(A) for the structure, then (B) and (C) for the bending). In this case, due to the geometry, bending to the left occurs when the left muscles are activated and then to the right for the symmetrical right muscles.

(A) Schematic diagram of C. elegans muscle fibers and its cross section, and it does not show the actin fibers. There are four muscle bands exist in the yellow layer. However, the yellow region is not an actual tissue layer and it is only used to define the position of the muscles. (B) Deformation diagram, left side muscles M1 and M2 are activated. (C) Deformation diagram, right side muscles M3 and M4 are activated. (D) Schematic diagram of C. elegans actin fibers and cross section. (E) Once the muscle is activated, the actin fiber orientation changes from the ‘loop’ to the ‘slope’, resulting in torque. (F) Schematic diagram of torsional and bending deformation.

To be more quantitative, we assume that the left side muscles are activated during a short period with an active constant strain value gm in the region M1 and M2, as shown in Fig.6(A)-(B); if the muscles are perfectly vertical, αm = βm = 0 in the initial configuration. In fact, the two muscles on the same side are not always completely in phase, and one may have a small delay. For simplicity, we assume that they contract simultaneously. During the entire initial period when the muscles are not activated, the actin fibers are distributed in a horizontal loop on the outer surface of the epidermis, but as soon as the muscle starts to contract, the actomyosin network will be reoriented (Lardennois et al. (2019)). The fibers are then distributed in an oblique pattern, which eventually causes the twisting of the embryo, see the schematic diagram in Fig.6(D)-(F). When this region is activated with a constant strain value, ga, the angle of the actin fibers will change according to the amplitude of the bending caused by the muscle contraction. In this situation, the angle of the actin fibers may change from , but βa is not changed and βa = 0.

Throughout the entire process, the muscle and actomyosin activities are assumed to operate simultaneously. Our modeling allows us to evaluate the bending and torsion generated independently by muscles and actin bundles, culminating in a complete deformation under coupling. Furthermore, the angle of the actomyosin fibers varies during muscle contraction. We maintain a constant activation of the actomyosin network and gradually increase the muscle activation.

As a result, the structure will be bent, causing a change in the angle of the actin fibers, ultimately yielding a deformation map that is presented in Fig.7(A)-(C). As myosin activation increased, we observed a consistent torsional deformation (Fig.7(E)) that agrees with the patterns seen in the video (Fig.7(D)). However, significant torsional deformations are not always present. In fact, other sources can induce torsion as a default symmetry of the muscle axis. We will now discuss torsion due to muscle contraction and the angle of deviation from the axis. In Fig.7(F)-(G), we demonstrate that the curvature provided by the model increases with increasing muscle activation increases and that the torsion is not simply related to the activation amplitude, as it also depends on the value of the angle αa, reaching a maximum at about π/4. Detailed calculations are given in Section Methods and Materials and Appendix 4.

Deformed configurations for different activation for muscles, (A) gm = −0.02, αa = π/3, ga = −0.01. (B) gm = −0.05, αa = π/4, ga = −0.01. (C) gm = −0.08, αa = π/6, ga = −0.01. (D) The graphs were captured from the Hymanlab, and the website: https://www.youtube.com/watch?v=M2ApXHhYbaw. The movie was acquired at a temperature of 200C using DIC optics. (E) gm = −0.1, αa = π/4, ga = −0.7. (F) Curvature is plotted as a function of muscle activation. (G) Torsion is plotted as a function of the actin activation and angle of actin fibers.

Energy transformation and Elongation

During the late elongation process, the four internal muscle bands contract cyclically in pairs (Yang (2017); Williams and Waterston (1994)). Each contraction of a pair increases the energy of the system under investigation, which is then rapidly released to the body. This energy exchange causes the torsional bending energy to be converted into elongation energy, leading to an increase in length during the relaxation phase, as shown in Fig.1 of the Appendix 5. With all the deformations obtained, Eq.(2) can be used to calculate the accumulated energy Wc produced by both the muscle and the actomyosin activities during a contraction. Subsequently, when the muscles on one side relax, the worm body returns to its original shape but with a tiny elongation corresponding to the transferred elastic energy. In this new state, the actin network assumes a ‘loop’ configuration with a strain of εga1 once relaxation is complete. If all the accumulated energy from the bending-torsion deformation is used to elongate the worm body, the accumulated energy Wc and the energy Wr after muscle relaxation are equivalent. The activation of actin fibers ga1 after muscle relaxation can be calculated and determined by our model.

Once the first phase is well characterized, we can quantify the total energy resulting from both muscle and actomyosin after each contraction. We then used our model to predict elongation in wild type C. elegans, unc-112(RNAi) mutant and spc-1(RNAi) pak-1(tm403) mutant and further compared the results with experimental observations. The wild-type length increases from about 90 μm to 210 μm during the muscle-activated phase (Lardennois et al. (2019)), the phase lasts about 140 minutes, and the average time interval between two contractions is about 40 seconds (Yang (2017)), so the number of contractions can be estimated to be about 210 times. Unfortunately, due to the difficulty of performing quantitative experiments on an embryo that is constantly moving in the egg shell between bending, torsion, and even rotations around its central axis, one can hypothesize this scenario: at each step i between the state Ai,0 to Ai,1 and then Ai,2, the whole mechanical muscle-myosin energy is transferred to the elongated step Ai,2, resulting a small δζi.

Considering the experimental results shown in Fig.8(B), we determine the optimal values for the activation parameters: gm = −0.15 and ga = −0.01 assuming that all the energy accumulated during the muscle activation is transferred to the elongation process (Wr = Wc). The elementary elongation δζi is gradually increased over time, which is shown as the black line in Fig.8(A). At the beginning, δζi is about 0.5μm, but at the end of this process, δζi is about 1.5μm, indicating that the worm will grow up to 290μm. The result is significantly larger than our actual size 210μm. As the elongation progresses, we assume that there is an energy transfer between bending-torsion-contraction and elongation but it may not be fully effective, meaning that a significant part of the energy is lost. From the experimental data, we estimate that the energy loss gradually increases, from full conversion at the beginning to only 40% of the accumulated energy used for elongation at the end of the process (Wr = 0.4Wc). For δζi it induces a first increase and then a decrease, which is shown as the blue line in Fig.8(A) and is responsible for the slowdown around 200 mins. This possibility, which may be not the only possible one, leads to the estimated elongation being in good agreement with the experimental data (see the blue-dashed curve in Fig.8(B)). Indeed it is possible that the C. elegans elongation requires other transformations that cost energy. As the embryo gradually elongates, energy dissipation and the biomechanical energy required to reorganize the actin bundles may be two factors contributing to the increased energy loss underlying the hypothesis.

(A) The elongation for each contraction varies with time. Black line: all energy converted to the elongation, blue line: partial energy converted to the elongation. The activation: gm = −0.15, ga = −0.01. (B) The model predicted results agree well with the experimental data of wild-type and different mutant C. elegans embryos (Lardennois et al. (2019)). Activation of the wild-type model (blue dashed line): gm = −0.15, ga = −0.01. The activation of unc-112(RNAi) (brown dashed line): gm = 0, ga = 0. In the case of pre-stretch failure (green dashed line), λ decreases from 1.8.

In addition, it has been reported in (Norman and Moerman (2002)) that the knockdown of unc-112(RNAi), known the to affect muscle contractions, leads to the arrest of elongation of the embryos at the twofold stage, indicating that the muscles have no activation, gm = 0 in our model, and no accumulated energy can be converted into elongation. Another mutation affecting the embryos consisting of mutant cells with pak-1(tm403), known to regulate the activity of myosin motors leads to a retraction of the embryo, so that the pre-stretch caused by myosin is not maintained and decreases. The results are fully consistent with a number of experimental observations and are shown in Fig.8(B).

Embryo rotations and dissipation

The main manifestation of the muscle activity, independent of the elongation, is probably the constant rotation of the embryo despite its confinement in the egg. This can be explained by a small angular deviation of the muscle sarcomeres from the central axis due to their attachment to the inner boundary of the cell epidermis, the so-called “dense bodies” (Moerman and Williams (2006)). Since they cross the horizontal plane at about ±45° and the deviation βm from the anterior-posterior axis is estimated to be about 6° each active muscle on the left (or on the right) contributes to the torque via a geometric factor of about ag = sin(6π/180) cos(π/4). Then a simple estimate of muscle activity in terms of torque is Λm ∼ μmπR3Smpmag(εgm) where Sm is the area of the left or right muscle pair compared to the section of the cylinder: Sm = 0.025 and pm is the distance of the muscles from the central axis of the embryo: pm = 0.75 while gm = 0.15 according to the elongation analysis. So the muscles on one side contribute to a torque Λm along the symmetry axis, given by Λm = 4.657μmπR3 · 10−5.

Let us now consider the dissipative torque, assuming that the rotational dynamics are stopped by friction after a bending event. Two cases can be considered: either the dissipation comes from the viscous flow or from the rubbing of the embryo when it folds. The fluid dissipation results from the rotations in the interstitial fluid inside and along the eggshell as the anterior-posterior axis remains parallel to the eggshell (Bhatnagar et al. (2023)) The interstitial fluid, of viscosity η contains a significant amount of sugars and other molecules necessary for embryo survival and that is more viscous than water (Labouesse (2023)). However, values for sucrose or sorbitol at the concentration of 1 mole/liter indicate a viscosity on the order of a few times the viscosity of water, which is 1 mPas. For example at 0.9 mole/liter and temperature of 20°, an aqueous solution of sorbitol has a viscosity of 1.6 mPas, which can be extrapolated to η = 1.9 mPas at 1.2 mole/liter (Jiang et al. (2013)). The estimate given by an embryo located in the middle of the egg shell, gives a weaker viscous torque once evaluated by according to a classical result reported by Landau et al. (Landau and Lifshitz (1971, 2013)).

It should be noted that this estimation assumes that the two cylinders: the egg shell and the embryo have the same axis of symmetry and concerns the beginning of the muscle activity where the radius is about 8.2 μm, the length is 90 μm, the radius of the shell is about Regg = 15μm and the length Legg = 54μm, see Fig.7(D). The angular velocity Ωe is more difficult to evaluate but it is about 90° per two seconds deduced from videos. As the embryo approaches the eggshell, the friction increases, and two eccentric cylinders of different radii must be considered with the two symmetry axes separated by a distance d. The hydrodynamic study in this case is far from being trivial, and seems to have been first initiated by Zhukoski (Zhukoski (1887)), who proposed the use of bipolar coordinates for the mathematical treatment. Many subsequent papers were published after, using barious simplified assumptions, and the study was completely revisited by Ballal and Rivlin (Ballal and Rivlin (1976)). Here we focus on the limit of a small gap δ between the rotating body shape and the egg and by considering an asymptotic analysis at small δ of the general result derived in (Ballal and Rivlin (1976)). Thus, the viscous torque reads where Lc is the zone of contact with the egg and d = Regg −δ −R. This approach is an approximation since the embryo is more toroidal than cylindrical (Yang Xinyi et al. (2023)), but the evaluation of the dissipation is satisfactory for δ = 0.5μm, Ωe ∼ π/4 S−1 and μ = 105P a. Going back back to the first model of dissipation with the same data, the ratio between the dissipative viscous torque and the active one gives: Λν/Λm = 0.02, which is obviously unsatisfactory. Finally, the dissipative energy ℰdiSS during one bending event leading to an angle of π/2 is ℰ = 1/2Λm ×(π/2)2 which represents 4% of the muscle elastic energy during the bending so at the beginning of the muscle activity (Appendix 5, Eq.(45)), the dissipation is present but is negligible. At the very end of the process, this ratio becomes 60% but as already mentioned, our estimate for the dissipation becomes very approximate, increases a lot due to the embryo confinement, and does not include the numerous biochemical steps necessary to reorganize the active network: actomyosin and muscles.

Discussion

Since the discovery of the pre-embryonic muscle activity in the C. elegans embryo, it has become critical to explain the role of mechanical forces generated by muscle contraction on the behavioral and functional aspects of the epidermis. We provide a mechanical model in which the C. elegans is simplified as a cylinder, and the muscle bands and actin that drive its elongation are modeled as active structures in a realistic position. We determine the fiber orientation from experimental observations and then calculate the deformation by tensorial analysis using the strains generated by the active components. Although a special focus is given to late elongation, its quantitative treatment cannot avoid the influence of the first stage of elongation due to the actomyosin network, which is responsible for a pre-strain of the embryo. In a finite elasticity formalism, the deformations induced by the muscles in a second step are coupled to the level of strains of the initial elongation period. For this, we must revisit the theory of the actomyosin contraction and previous results (Ciarletta et al. (2009); Vuong-Brender et al. (2017a); Ben Amar et al. (2018)) to unify the complete treatment. In particular, a model for the recruitment of active myosin motors under forcing is presented that recovers the experimental results of the first elongation phase

The elongation process of C. elegans during the late period is much more complex than the early elongation stage which is caused by actin contraction alone. During late elongation, the worm is deformed by the combined action of muscle and actomyosin, resulting in an energy-accumulating process. Bending deformation is a phenomenon resulting from unilateral muscle contraction, and during the late elongation, significant torsional deformation is observed, indicating that the bending process induces a reorientation of the actin fibers. It should be noted that the embryo always rotates in parallel to the muscle activity, which makes any experimental measurements difficult. However, our model can predict that if the muscles are not perfectly vertical,a torque is generated that causes rotation and eventually torsion. The accumulated energy is then partially converted into energy for the ongoing action of actin, allowing the embryo to elongate when the muscle relaxes. Both sides of the C. elegans muscles contract in a sequential cycle, repeating the energy conversion process, and eventually completing the elongation process. However, the energy exchange between bending and elongation is limited, among other factors, by the viscous dissipation induced by rotation, which is also evaluated in this study. The necessary reorganization of the active networks (actomyosin and muscles) due to this enormous shape change of the embryo is not investigated here in detail. Parallel to the elongation, the cuticle is built around the body (Page and Johnstone (2007)). This very thin and stiff membrane provides protection and locomotion after hatching. Obviously, these processes will interfere with muscle activity. These two aspects, which intervene in the final stage of the worm’s confinement, play a very important role at the interface between genetics, biochemistry, and mechanics.

Finally, the framework presented here not only provides a theoretical explanation for embryonic elongation in C. elegans, but can also be used to model other biological behaviors, such as plant tropism (Moulton et al. (2020b)) and elephant trunk elongation (Schulz et al. (2022b,a)) and bending. Our ideas could potentially be used in the emerging field of soft robotics, such as octopus leg inspired robots (Kang et al. (2012); Nakajima et al. (2013); Calisti et al. (2015)), which are soft and their deformation is induced by muscle activation. We can reliably predict the deformation by knowing the position of activation and the magnitude of fthe orces in the model. In addition, residual stresses can be incorporated into our model to achieve design goals.

Methods and Materials

The model has been presented in a series of articles by A. Goriely and collaborators (Goriely and Tabor (2013); Moulton et al. (2013, 2020b); Kaczmarski et al. (2022); Goriely et al. (2022)). As one can imagine, this is far from trivial, and most of the literature on the subject concerns either full cylinders or cylindrical shells in torsion around the axis of symmetry, the case of bending is far from trivial. However, the geometry of the muscles in C. elegans automatically leads to a bending process that cannot be discarded. We take advantage of these previous works and apply the methodology to our model.

The deformation gradient

The finite strain G0 maps the initial stress-free state ℬ0 to a state ℬ1, which describes the early elongation. We then impose an incremental strain filed G1 that maps the state ℬ1 to the state ℬ2, which represents late elongation. So, the deformation gradient can be expressed as: F = FeG1G0 (Goriely and Ben Amar (2007)). The deformation gradient F = Fij di ⊗ ej, i ∈ {1, 2, 3} and j labels the reference coordinates {R, Θ, Z} finally reads: where λ is the axial extension due to the pre-strained.

We define G0 = Diag (1, 1 + εc, 1), since g0 = 0.88 can be determined by the early elongation, we write it in the form of 1 + εc to simplify the calculation, c ≈ −0.6 (ε ≈ 0.2).

Consider fibers oriented along the unit vector m = (sin α sin β, sin α cos β, cos α) for the incremental strain G1, where α and β characterize the angles with ez and eθ and can be found in Fig.3(A). The active filamentary tensor in cylindrical coordinates is then G = G0 (I + εg m ⊗ m), see (Holzapfel (2000)) which reads:

The energy function

The associated strain-energy density can be obtained by expanding the inner variables ai (see Eq.(1)), pi and the potential V (see Eq.(2)) as: For each cross section, the associated Euler-Lagrange equations take the following form: where j = 1, 2, 3, k = 0, 1, in association with boundary conditions at each order k = 0, 1 on the outer radius Ri: We solve these equations order by order, taking incompressibility into account. At the lowest order, the incompressibility imposes the deformation: a(0) = (r (R), Θ, 0) and the Euler-Lagrange equation gives the Lagrange parameter p(0) = P0(R) which finally read: and the boundary condition σrr(R = 1) = 0 gives the value of P0 = 1/λ. At 𝒪 (ε), the Euler-Lagrange equations are again automatically satisfied, and at 𝒪 (ε2), the crucial question is to get the correct expression for and p(1). Based on the previous subsection, the following form is intuited: As before, the Euler-Lagrange equations and the incompressibility condition give the two constants and h1, h2, h3 are a function of R. With the solutions for active strain a(0) and a(1), full expressions of them are given in the Appendix 4. The second order energy takes the form: and we can transform it into this form: where Ai, Bi, Ci and Di are functions of R and Θ.

The method for determining deformation

Comparing with the energy of an extensible elastic rod (Moulton et al. (2020a); Kirchhoff and Hensel (1883); Mielke (1988); Mora and Müller (2002, 2004)), we recognize the classic extensional stiffness K0, bending stiffness K1 and K2, torsional stiffness K3 coefficients: where A2, B2, C2 and D2 are related to the shear modulus μ, so for a uniform material with no variations of shear modulus, SK represents the cross section of the cylinder. But if not, we need to divide different regions to perform the integration.

We now focus on the intrinsic extension and curvature of the cylindrical object induced by the active strains, which requires the competition of the active forces with the stiffness of the cylindrical beam according to the relationships, where Hi calculated by: where A1, B1, C1 and D1 are related to the shear modulus μ,the fiber angles α and β, and the activation g. So, the integration region SH is divided into the different parts of the embryo which all contribute to the deformation.

To calculate the intrinsic Frenet curvature and torsion, we use the following equation: Finally, in the case of the activated left side muscles, we can calculate the intrinsic extension and curvatures using Eq.16), û1m = 0 and û3m = 0, so Frenet’s intrinsic curvature and torsion Eq.(18), and . For the activated actin case, we can calculate the intrinsic extension and curvatures by Eq.(16), û1a = 0 and û2a = 0, and the Frenet’s intrinsic curvature and torsion Eq.(18), and . These quantities have been used to calculate the deformation images, as shown previously in Fig.7.

After obtaining all solutions, we can use these quantities to calculate the accumulated energy Wc in the system after one contraction by Eq.(2). By defining the activation (gm and ga) and the energy conversion efficiency, we can obtain the activation (ga1) of the actin bundles after a single contraction and further calculate the elongation.

All calculation results and parameters are presented in the Appendix.